Code by Johanna Schönecker

Edited 9th July 2025

This quarto document contains code to load and pre-process pixel-, fire- and patch-level data for megafires in the Sierra Nevada ecoregion which burned between 1985 and 2023.

Different sections contain code to produce figures.

##Packages

```{r Loading packages}
# -------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyr, raster, rgdal, ggplot2, ggspatial, sp, sf, rgeos, reshape,
  patchwork, Rcpp, scales, dplyr, ggpubr, landscapemetrics, landscapetools,
  here, stringr, moments, reshape2, ggeasy, vctrs, mgcv, networkD3,
  ggridges, cowplot, factoextra, gdistance, fs, data.table, terra,
  gridExtra, grid, car, biscale, e1071
)

```

##Data loading and prep

```{r}
#Fire level datasets
fire_metrics_mega <- fread(paste0(here(),"/Data/SN_megafires_1985_2023.csv"), header=TRUE)

#Patch level datasets- rdnbr
patch_metrics_mega <- fread(paste0(here(),"/Data/SN_patch_metrics_mega_rdnbr_1985_2023.csv"))

megafire_IDs <- c(5529, 5718, 5721, 8026, 7808, 7747, 7996, 7961, 
                   7412, 7514, 7242, 9714, 9724, 9487, 8985, 9039, 
                   8747, 10495, 10498, 10504, 10094, 10273, 1819, 1772, 
                   2111, 2140, 1210, 1211, 1216, 1215, 1220, 1319, 
                   1315, 718, 911, 932, 869, 888, 945, 959, 504, 
                   3363, 3567, 3573, 3574, 3051, 3094, 3188, 2706, 
                   2147, 2531, 2415, 2618, 5207, 5309, 4277, 4285, 
                   4306, 4380, 4386, 4374, 4500, 3983, 3992)
```

##Megafire temporal trends

#Figure 1

```{r}
# -------------------------------------------------------------------------
# Find out area burned annually in normal & megafires
# -------------------------------------------------------------------------

SN_fires_attr_all <- fread(paste0(here(), "/Data/SN_all_fires_1985_2023.csv"), header = TRUE)
SN_fires_attr_mega <- fire_metrics_mega

# Summarize total area burned per year
area_burned_all_annual <- SN_fires_attr_all %>%
  group_by(YEAR_) %>%
  summarise(total_burned_area = sum(Shape_Area, na.rm = TRUE))

# Summarize total megafire area per year
area_burned_mega_annual <- SN_fires_attr_mega %>%
  group_by(YEAR_) %>%
  summarise(mega_area = sum(Shape_Area, na.rm = TRUE))

# Combine both summaries
area_burned_combined <- area_burned_all_annual %>%
  left_join(area_burned_mega_annual, by = "YEAR_") %>%
  mutate(mega_area = ifelse(is.na(mega_area), 0, mega_area))

# Calculate small fire area
area_burned_combined$small_area <- area_burned_combined$total_burned_area - area_burned_combined$mega_area

# -------------------------------------------------------------------------
# Megafire counts and total burned area
# -------------------------------------------------------------------------

mega_counts_area <- data.frame(
  year = c(
    1985, 1986, 1987, 1988, 1989, 1990, 1991, 1992, 1993, 1994,
    1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004,
    2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014,
    2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023
  ),
  count_mega = c(
    0, 0, 4, 1, 0, 2, 0, 1, 0, 2, 0, 1, 0, 0, 3, 2, 1, 2, 0, 0,
    0, 0, 2, 3, 0, 0, 1, 6, 2, 3, 2, 2, 4, 4, 1, 7, 7, 1, 0
  ),
  area_burned = area_burned_combined$total_burned_area
)

mega_counts_area$area_burned_ha <- mega_counts_area$area_burned * 0.0001

# -------------------------------------------------------------------------
# Linear model: area burned vs. number of megafires
# -------------------------------------------------------------------------

model <- lm(area_burned_ha ~ count_mega, data = mega_counts_area)

# Predictions + confidence intervals
mega_counts_area$predictions <- predict(model, newdata = mega_counts_area, type = "response")
confidence_data <- predict(model, newdata = mega_counts_area, interval = "confidence", level = 0.95)

mega_counts_area$lower_bound <- confidence_data[, "lwr"]
mega_counts_area$upper_bound <- confidence_data[, "upr"]

# -------------------------------------------------------------------------
# Convert wide to long for stacked area plot
# -------------------------------------------------------------------------

area_burned_long <- area_burned_combined %>%
  dplyr::select(-total_burned_area) %>%
  pivot_longer(
    cols = c(mega_area, small_area),
    names_to = "size_class",
    values_to = "area"
  ) %>%
  mutate(size_class = ifelse(size_class == "mega_area", "mega", "small"))

# -------------------------------------------------------------------------
# Plot: relationship between number of megafires and total area burned
# -------------------------------------------------------------------------

plot_31 <- ggplot(mega_counts_area, aes(x = count_mega, y = area_burned_ha)) +
  geom_point(aes(color = "Data"), alpha = 1.0, size = 3.5) +
  geom_line(aes(y = predictions, color = "Fitted"), size = 1) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "darkgrey", alpha = 0.3) +
  scale_color_manual(values = c("Data" = "black", "Fitted" = "black")) +
  labs(x = "Number of megafires in a fire year", y = "Area burned (×1000 ha)") +
  scale_y_continuous(labels = scales::label_number(scale = 1 / 1000, suffix = "", accuracy = 1)) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(color = "black", size = 18),
    axis.title.x = element_text(color = "black", size = 18),
    axis.text = element_text(size = 14, face = "bold")
  )

# Add R² annotation
rsq <- summary(model)$r.squared
plot_31 <- plot_31 +
  annotate(
    "text",
    x = max(mega_counts_area$count_mega),
    y = min(mega_counts_area$area_burned_ha),
    label = paste("R =", round(rsq, 4)),
    hjust = 1, vjust = -1,
    color = "black", size = 8
  )

print(plot_31)

# -------------------------------------------------------------------------
# Plot: annual area burned by fire size class and number of megafires
# -------------------------------------------------------------------------

area_burned_long <- area_burned_combined %>%
  pivot_longer(
    cols = c(mega_area, small_area),
    names_to = "type",
    values_to = "area"
  ) %>%
  mutate(area_ha = area / 10000)

plot_23_1 <- ggplot() +
  geom_bar(
    data = area_burned_long,
    aes(x = YEAR_, y = area_ha, fill = type),
    stat = "identity"
  ) +
  geom_point(
    data = mega_counts_area,
    aes(x = year, y = count_mega * 100000),
    size = 3, color = "black", alpha = 1.0
  ) +
  geom_smooth(
    data = mega_counts_area,
    aes(x = year, y = count_mega * 100000),
    method = "glm",
    method.args = list(family = "poisson"),
    se = TRUE, color = "black"
  ) +
  scale_x_continuous(name = "Fire year", limits = c(1984, 2023)) +
  scale_y_continuous(
    name = "Area burned (×1000 ha)",
    labels = scales::label_number(scale = 1 / 1000, suffix = "", accuracy = 1),
    limits = c(0, 700000),
    sec.axis = sec_axis(~ . / 100000, name = "Number of megafires")
  ) +
  scale_fill_manual(values = c("small_area" = "#E17327", "mega_area" = "#932191")) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(color = "black", size = 18),
    axis.title.y.right = element_text(color = "black", size = 18),
    axis.title.x = element_text(color = "black", size = 18),
    axis.text = element_text(size = 14, face = "bold"),
    legend.position = "none"
  ) +
  labs(fill = "Fire Size")

print(plot_23_1)

# -------------------------------------------------------------------------
# Combine both plots into one figure
# -------------------------------------------------------------------------

library(cowplot)

figure_1 <- plot_grid(
  plot_23_1, plot_31,
  ncol = 1,
  align = "v",
  labels = c("b", "c"),
  label_size = 16
)

print(figure_1)

```

##Patch area/ size (cumulative)

```{r}
# -------------------------------------------------------------------
# Patch sizes and area by patch size class, all patches
# -------------------------------------------------------------------

# Bin 'n' and prepare dataset
patch_metrics_mega <- patch_metrics_mega %>%
  mutate(
    n_bins = cut(
      n,
      breaks = c(1, 100, 1000, 10000, 1500000),
      labels = c("small", "medium", "large", "very large"),
      include.lowest = TRUE
    )
  )

patch_metrics_mega$area_ha       <- patch_metrics_mega$n * 0.09
patch_metrics_mega$outer_area_ha <- patch_metrics_mega$area_ha - patch_metrics_mega$core_area

# Step 1: Pivot to long format for stacked bars
patch_long <- patch_metrics_mega %>%
  pivot_longer(
    cols      = c(core_area, outer_area_ha),
    names_to  = "area_type",
    values_to = "area_val"
  )

# Step 2: Summarise total area per bin and type
patch_summary <- patch_long %>%
  group_by(n_bins, area_type) %>%
  summarise(total_area = sum(area_val, na.rm = TRUE), .groups = "drop")

# Step 3: Totals, counts, and cumulative areas per bin
patch_counts <- patch_metrics_mega %>%
  group_by(n_bins) %>%
  summarise(count = n(), .groups = "drop")

patch_totals <- patch_summary %>%
  group_by(n_bins) %>%
  summarise(total_area_sum = sum(total_area), .groups = "drop") %>%
  left_join(patch_counts, by = "n_bins") %>%
  arrange(as.numeric(n_bins)) %>%
  mutate(cumulative_area = cumsum(total_area_sum))

patch_core_cumulative <- patch_summary %>%
  filter(area_type == "core_area") %>%
  arrange(as.numeric(n_bins)) %>%
  group_by(n_bins) %>%
  summarise(core_area_total = sum(total_area, na.rm = TRUE), .groups = "drop") %>%
  mutate(cumulative_core_area = cumsum(core_area_total))

patch_outer_cumulative <- patch_summary %>%
  filter(area_type == "outer_area_ha") %>%
  arrange(as.numeric(n_bins)) %>%
  group_by(n_bins) %>%
  summarise(outer_area_total = sum(total_area, na.rm = TRUE), .groups = "drop") %>%
  mutate(cumulative_outer_area = cumsum(outer_area_total))

# Step 4: Plot
cum_1 <- ggplot(patch_summary, aes(x = as.factor(n_bins), y = total_area, fill = area_type)) +
  geom_bar(stat = "identity") +

  # Count labels
  geom_text(
    data = patch_totals,
    aes(x = as.factor(n_bins), y = total_area_sum, label = scales::comma(count)),
    vjust = -0.5, size = 3.5, inherit.aes = FALSE
  ) +

  # Cumulative total area line
  geom_line(
    data = patch_totals,
    aes(x = as.factor(n_bins), y = cumulative_area, group = 1),
    color = "darkred", size = 1, inherit.aes = FALSE
  ) +
  geom_point(
    data = patch_totals,
    aes(x = as.factor(n_bins), y = cumulative_area),
    color = "darkred", size = 2, inherit.aes = FALSE
  ) +

  # Fill & labels
  scale_fill_manual(
    values = c("core_area" = "orange", "outer_area_ha" = "lightgrey"),
    labels = c("core_area" = "core area", "outer_area_ha" = "outer area"),
    name   = NULL
  ) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "",
    x = "Patch size",
    y = "Total area (ha)",
    fill = "Area type"
  ) +
  theme_classic() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x      = element_text(angle = 0, hjust = 1),
    axis.title       = element_text(face = "bold"),
    axis.text        = element_text(face = "bold")
  ) +

  # Cumulative core area (dashed)
  geom_line(
    data = patch_core_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_core_area, group = 1),
    color = "darkorange", size = 1, linetype = "dashed", inherit.aes = FALSE
  ) +
  geom_point(
    data = patch_core_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_core_area),
    color = "darkorange", size = 2, inherit.aes = FALSE
  ) +

  # Cumulative outer area (dashed)
  geom_line(
    data = patch_outer_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_outer_area, group = 1),
    color = "grey40", size = 1, linetype = "dashed", inherit.aes = FALSE
  ) +
  geom_point(
    data = patch_outer_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_outer_area),
    color = "grey40", size = 2, inherit.aes = FALSE
  ) +

  # Annotations
  annotate(
    "text",
    x = Inf, y = max(patch_totals$cumulative_area, na.rm = TRUE),
    label = "cumulative total area", color = "darkred",
    hjust = 1.1, vjust = -0.5, size = 3.5
  ) +
  annotate(
    "text",
    x = Inf, y = max(patch_core_cumulative$cumulative_core_area, na.rm = TRUE),
    label = "cumulative core area", color = "darkorange",
    hjust = 1.1, vjust = -0.5, size = 3.5
  ) +
  annotate(
    "text",
    x = Inf, y = max(patch_outer_cumulative$cumulative_outer_area, na.rm = TRUE),
    label = "cumulative outer area", color = "grey40",
    hjust = 1.1, vjust = -0.5, size = 3.5
  )

```

##Patch area/ size high sev (cumulative)

```{r}
# -------------------------------------------------------------------
# Only high severity
# Patch sizes
# -------------------------------------------------------------------

# Binning 'n' and preparing the dataset
patch_metrics_mega <- patch_metrics_mega %>%
  mutate(
    n_bins = cut(
      n,
      breaks = c(1, 100, 1000, 10000, 1500000),
      labels = c("small", "medium", "large", "very large"),
      include.lowest = TRUE
    )
  )

patch_metrics_mega$area_ha       <- patch_metrics_mega$n * 0.09
patch_metrics_mega$outer_area_ha <- patch_metrics_mega$area_ha - patch_metrics_mega$core_area

patch_metrics_mega_sev <- patch_metrics_mega %>%
  filter(rdnbr_class == 3)

# Step 1: Pivot to long format for stacked bars
patch_long <- patch_metrics_mega_sev %>%
  pivot_longer(
    cols      = c(core_area, outer_area_ha),
    names_to  = "area_type",
    values_to = "area_val"      # ← renamed from area_ha
  )

# Step 2: Summarise total area and patch count per n_bins
patch_summary <- patch_long %>%
  group_by(n_bins, area_type) %>%
  summarise(total_area = sum(area_val, na.rm = TRUE), .groups = "drop")

# Step 3: Calculate cumulative total area and counts per bin (regardless of type)
patch_counts <- patch_metrics_mega_sev %>%
  group_by(n_bins) %>%
  summarise(count = n(), .groups = "drop")

patch_totals <- patch_summary %>%
  group_by(n_bins) %>%
  summarise(total_area_sum = sum(total_area), .groups = "drop") %>%
  left_join(patch_counts, by = "n_bins") %>%
  arrange(as.numeric(n_bins)) %>%
  mutate(cumulative_area = cumsum(total_area_sum))

patch_core_cumulative <- patch_summary %>%
  filter(area_type == "core_area") %>%
  arrange(as.numeric(n_bins)) %>%
  group_by(n_bins) %>%
  summarise(core_area_total = sum(total_area, na.rm = TRUE), .groups = "drop") %>%
  mutate(cumulative_core_area = cumsum(core_area_total))

patch_outer_cumulative <- patch_summary %>%
  filter(area_type == "outer_area_ha") %>%
  arrange(as.numeric(n_bins)) %>%
  group_by(n_bins) %>%
  summarise(outer_area_total = sum(total_area, na.rm = TRUE), .groups = "drop") %>%
  mutate(cumulative_outer_area = cumsum(outer_area_total))

# Step 4: Plot
cum_2 <- ggplot(patch_summary, aes(x = as.factor(n_bins), y = total_area, fill = area_type)) +
  geom_bar(stat = "identity") +

  # Add count labels (from patch_totals)
  geom_text(
    data = patch_totals,
    aes(x = as.factor(n_bins), y = total_area_sum, label = comma(count)),
    vjust = -0.5, size = 3.5, inherit.aes = FALSE
  ) +

  # Add cumulative area line (also from patch_totals)
  geom_line(
    data = patch_totals,
    aes(x = as.factor(n_bins), y = cumulative_area, group = 1),  # ← only aesthetics in patch_totals
    color = "darkred", size = 1, inherit.aes = FALSE
  ) +
  geom_point(
    data = patch_totals,
    aes(x = as.factor(n_bins), y = cumulative_area),
    color = "darkred", size = 2, inherit.aes = FALSE
  ) +

  scale_fill_manual(
    values = c("core_area" = "orange", "outer_area_ha" = "lightgrey"),
    labels = c("core_area" = "core area", "outer_area_ha" = "outer area"),
    name   = NULL  # This removes the title
  ) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "",
    x     = "Patch size",
    y     = "Total area (ha)",
    fill  = "Area type"
  ) +
  theme_classic() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x      = element_text(angle = 0, hjust = 1)
  ) +

  # Add second cumulative line for core area
  geom_line(
    data = patch_core_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_core_area, group = 1),
    color = "darkorange", size = 1, inherit.aes = FALSE, linetype = "dashed"
  ) +
  geom_point(
    data = patch_core_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_core_area),
    color = "darkorange", size = 2, inherit.aes = FALSE
  ) +

  # Add cumulative line for outer area
  geom_line(
    data = patch_outer_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_outer_area, group = 1),
    color = "grey40", size = 1, inherit.aes = FALSE, linetype = "dashed"
  ) +
  geom_point(
    data = patch_outer_cumulative,
    aes(x = as.factor(n_bins), y = cumulative_outer_area),
    color = "grey40", size = 2, inherit.aes = FALSE
  ) +

  annotate(
    "text",
    x = Inf, y = max(patch_totals$cumulative_area, na.rm = TRUE),
    label = "cumulative total area", color = "darkred",
    hjust = 1.1, vjust = -0.5, size = 3.5
  ) +
  annotate(
    "text",
    x = Inf, y = max(patch_core_cumulative$cumulative_core_area, na.rm = TRUE),
    label = "cumulative core area", color = "darkorange",
    hjust = 1.1, vjust = -0.5, size = 3.5
  ) +
  annotate(
    "text",
    x = Inf, y = max(patch_outer_cumulative$cumulative_outer_area, na.rm = TRUE),
    label = "cumulative outer area", color = "grey40",
    hjust = 1.1, vjust = -0.5, size = 3.5
  )

```

#AWMPS- area weighted mean patch size

```{r}
# -------------------------------------------------------------------
# Function to calculate AWMPS for a given set of patch areas
# -------------------------------------------------------------------

calculate_awmps <- function(patch_areas) {
  total_area <- sum(patch_areas)
  awmps <- sum(patch_areas^2) / total_area
  return(awmps)
}

# -------------------------------------------------------------------
# Create empty dataframe to store results
# -------------------------------------------------------------------
patch_size_aw_mega <- data.frame(
  rdnbr_class = integer(),
  AWMPS       = numeric(),
  OBJECTID    = integer()
)

# -------------------------------------------------------------------
# Loop through each megafire ID and calculate AWMPS
# -------------------------------------------------------------------
for (i in megafire_IDs) {
  
  patch_metrics_indiv <- patch_metrics_mega %>%
    filter(OBJECTID == i)
  
  # Calculate AWMPS for each severity class
  awmps_by_class <- patch_metrics_indiv %>%
    group_by(rdnbr_class) %>%
    summarise(AWMPS = calculate_awmps(n), .groups = "drop")
  
  awmps_by_class$OBJECTID <- i
  
  patch_size_aw_mega <- rbind(patch_size_aw_mega, awmps_by_class)
}

# -------------------------------------------------------------------
# Clean and recode severity classes
# -------------------------------------------------------------------
patch_size_aw_mega <- patch_size_aw_mega %>%
  filter(rdnbr_class != -1) %>%
  mutate(
    rdnbr_class = as.character(rdnbr_class),
    rdnbr_class = dplyr::recode(
      rdnbr_class,
      "1" = "low",
      "2" = "moderate",
      "3" = "severe"
    ),
    awmps_ha = AWMPS * 0.09
  )


```

# LPI

```{r}
# -------------------------------------------------------------------
# Calculate the LPI for every class in every megafire
# -------------------------------------------------------------------

# Path to classified RdNBR rasters
raster_dir <- paste0(here(),"/Data/Rasters/RdNBR_classified")

# List all .tif raster files in the folder
raster_files <- fs::dir_ls(raster_dir, regexp = "\\.tif$")

# Function to calculate LPI per class for a raster
calculate_lpi <- function(file_path) {
  # Load raster
  rast_obj <- terra::rast(file_path)

  # Convert to RasterLayer (landscapemetrics prefers Raster*)
  rast_r <- raster::raster(rast_obj)

  # Calculate LPI for each class
  lpi_result <- landscapemetrics::lsm_c_lpi(rast_r)

  # Add raster name to identify the source
  lpi_result$raster_name <- basename(file_path)

  lpi_result
}

# Apply to all rasters and combine
lpi_list <- lapply(raster_files, calculate_lpi)
lpi_df   <- dplyr::bind_rows(lpi_list)

# -------------------------------------------------------------------
# Prepare a wide table for supplement
# -------------------------------------------------------------------

# Step 1: Extract fire ID from raster_name and join fire names
lpi_df <- lpi_df %>%
  dplyr::mutate(
    OBJECTID = as.numeric(stringr::str_remove(raster_name, "\\.tif$"))
  ) %>%
  dplyr::left_join(
    fire_metrics_mega %>% dplyr::select(OBJECTID, FIRE_NAME),
    by = "OBJECTID"
  )

# Step 2: Recode class into labels
lpi_df <- lpi_df %>%
  dplyr::mutate(
    class_label = dplyr::case_when(
      class == 0 ~ "Unchanged",
      class == 1 ~ "Low",
      class == 2 ~ "Moderate",
      class == 3 ~ "High",
      TRUE       ~ as.character(class)
    )
  )

# Step 3: Summarise to one row per fire with total value per class
lpi_summary <- lpi_df %>%
  dplyr::group_by(FIRE_NAME, class_label) %>%
  dplyr::summarise(total_value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from  = class_label,
    values_from = total_value
  ) %>%
  dplyr::rename(`Fire name` = FIRE_NAME) %>%
  dplyr::select(`Fire name`, Unchanged, Low, Moderate, High) %>%
  dplyr::left_join(
    fire_metrics_mega %>% dplyr::select(FIRE_NAME, YEAR_),
    by = c("Fire name" = "FIRE_NAME")
  ) %>%
  dplyr::rename(`Fire year` = YEAR_) %>%
  dplyr::select(`Fire name`, `Fire year`, Unchanged, Low, Moderate, High)

# Write CSV
data.table::fwrite(
  lpi_summary, paste0(here(),"/Outputs/LPI_all_megafires.csv")
)


```

# Figure 3: Patch characteristics

```{r}
# -------------------------------------------------------------------
# Data preparation
# -------------------------------------------------------------------
# Bin patch sizes and drop very small patches (n <= 2)
patch_data <- patch_metrics_mega %>%
  dplyr::mutate(
    n_bins = cut(
      n,
      breaks = c(1, 100, 1000, 10000, 1500000),
      labels = c("small", "medium", "large", "very large"),
      include.lowest = TRUE
    )
  ) %>%
  dplyr::filter(n > 2)

# Optional filter for conifer vegetation
# patch_data <- patch_data %>% dplyr::filter(mode_veg == 7)

# -------------------------------------------------------------------
# Create metrics
# -------------------------------------------------------------------
# Total core area per patch
core_area_summary <- patch_data %>%
  dplyr::group_by(OBJECTID, n_bins, rdnbr_class) %>%
  dplyr::summarise(core_area_sum = sum(core_area), .groups = "drop")

# Contiguity mean
contiguity_summary <- patch_data %>%
  dplyr::group_by(OBJECTID, n_bins, rdnbr_class) %>%
  dplyr::summarise(contiguity_mean = mean(contiguity), .groups = "drop")

# Perimeter-area ratio mean
paratio_summary <- patch_data %>%
  dplyr::group_by(OBJECTID, n_bins, rdnbr_class) %>%
  dplyr::summarise(perimeter_area_mean = mean(perimeter_area), .groups = "drop")

# -------------------------------------------------------------------
# Harmonize RdNBR class labels
# -------------------------------------------------------------------
recode_rdnbr <- function(df) {
  df %>%
    dplyr::filter(rdnbr_class %in% c(1, 2, 3)) %>%
    dplyr::mutate(
      rdnbr_class = dplyr::recode(
        as.character(rdnbr_class),
        "1" = "low",
        "2" = "moderate",
        "3" = "high"
      ),
      rdnbr_class = factor(rdnbr_class, levels = c("low", "moderate", "high"))
    )
}

core_area_burned  <- recode_rdnbr(core_area_summary)
contiguity_burned <- recode_rdnbr(contiguity_summary)
paratio_burned    <- recode_rdnbr(paratio_summary)

# -------------------------------------------------------------------
# Boxplots of patch metrics by patch size
# -------------------------------------------------------------------
# Colors
rdnbr_colors_named <- c("low" = "#F6BE00", "moderate" = "#21918c", "high" = "#440154")
rdnbr_colors       <- c("1"   = "#F6BE00", "2"        = "#21918c", "3"   = "#440154")

# Common theme
box_theme <- ggplot2::theme_classic() +
  ggplot2::theme(
    strip.background = ggplot2::element_rect(fill = "lightgrey", color = "grey"),
    strip.text       = ggplot2::element_text(color = "black"),
    axis.text        = ggplot2::element_text(face = "bold", size = 18),
    axis.title       = ggplot2::element_text(size = 20),
    legend.position  = "none"
  )

plot_core_area <- ggplot2::ggplot(
  core_area_burned,
  ggplot2::aes(x = n_bins, y = core_area_sum, fill = rdnbr_class)
) +
  ggplot2::geom_boxplot(alpha = 0.5) +
  ggplot2::scale_fill_manual(values = rdnbr_colors_named) +
  ggplot2::labs(x = "Patch size class", y = "Core area within patch (ha)") +
  ggplot2::coord_cartesian(ylim = c(0, 7500)) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  box_theme

plot_paratio <- ggplot2::ggplot(
  paratio_burned,
  ggplot2::aes(x = n_bins, y = perimeter_area_mean, fill = rdnbr_class)
) +
  ggplot2::geom_boxplot(alpha = 0.5) +
  ggplot2::scale_fill_manual(values = rdnbr_colors_named) +
  ggplot2::labs(x = "Patch size class", y = "Perimeter:area") +
  box_theme

plot_contiguity <- ggplot2::ggplot(
  contiguity_burned,
  ggplot2::aes(x = n_bins, y = contiguity_mean, fill = rdnbr_class)
) +
  ggplot2::geom_boxplot(alpha = 0.5) +
  ggplot2::scale_fill_manual(values = rdnbr_colors_named) +
  ggplot2::labs(x = "Patch size class", y = "Contiguity") +
  box_theme

# -------------------------------------------------------------------
# Patch metrics over time (mean per fire year)
# -------------------------------------------------------------------
patch_data <- patch_data %>%
  dplyr::mutate(rdnbr_class = factor(rdnbr_class))

summary_by_year <- patch_data %>%
  dplyr::group_by(OBJECTID, rdnbr_class) %>%
  dplyr::summarise(
    mean_core_area     = mean(core_area, na.rm = TRUE),
    mean_perimeter_area = mean(perimeter_area, na.rm = TRUE),
    mean_contiguity    = mean(contiguity, na.rm = TRUE),
    fire_year          = mean(fire_year, na.rm = TRUE),
    .groups            = "drop"
  )

line_theme <- ggplot2::theme_classic() +
  ggplot2::theme(
    axis.text       = ggplot2::element_text(face = "bold"),
    axis.title      = ggplot2::element_text(face = "bold"),
    legend.position = "none"
  )

lineplot_core_area <- ggplot2::ggplot(
  summary_by_year,
  ggplot2::aes(x = fire_year, y = mean_core_area, color = rdnbr_class)
) +
  ggplot2::geom_point(size = 2.5, alpha = 0.7) +
  ggplot2::geom_smooth(method = "gam", se = TRUE, alpha = 0.2) +
  ggplot2::scale_color_manual(values = rdnbr_colors) +
  ggplot2::labs(y = "Mean core area within patch (ha)", x = "Fire year") +
  line_theme

lineplot_paratio <- ggplot2::ggplot(
  summary_by_year,
  ggplot2::aes(x = fire_year, y = mean_perimeter_area, color = rdnbr_class)
) +
  ggplot2::geom_point(size = 2.5, alpha = 0.7) +
  ggplot2::geom_smooth(method = "gam", se = TRUE, alpha = 0.2) +
  ggplot2::scale_color_manual(values = rdnbr_colors) +
  ggplot2::labs(y = "Perimeter:area", x = "Fire year") +
  line_theme

lineplot_contiguity <- ggplot2::ggplot(
  summary_by_year,
  ggplot2::aes(x = fire_year, y = mean_contiguity, color = rdnbr_class)
) +
  ggplot2::geom_point(size = 2.5, alpha = 0.7) +
  ggplot2::geom_smooth(method = "gam", se = TRUE, alpha = 0.2) +
  ggplot2::scale_color_manual(values = rdnbr_colors) +
  ggplot2::labs(y = "Contiguity", x = "Fire year") +
  line_theme

# -------------------------------------------------------------------
# Combine plots
# -------------------------------------------------------------------
# Top row (boxplots)
top_combined <- cowplot::plot_grid(
  plot_paratio     + ggplot2::theme(axis.title.x = ggplot2::element_blank()),
  plot_contiguity  + ggplot2::theme(axis.title.x = ggplot2::element_blank()),
  plot_core_area   + ggplot2::theme(axis.title.x = ggplot2::element_blank()),
  labels    = c("A", "B", "C"),
  label_size = 20,
  ncol      = 3
)

# Bottom row (time trends)
bottom_combined <- cowplot::plot_grid(
  lineplot_paratio     + ggplot2::theme(axis.title.x = ggplot2::element_blank()),
  lineplot_contiguity  + ggplot2::theme(axis.title.x = ggplot2::element_blank()),
  lineplot_core_area   + ggplot2::theme(axis.title.x = ggplot2::element_blank()),
  labels = c("A", "B", "C"),
  ncol   = 3
)

# Shared x-axis labels
top_labeled <- cowplot::plot_grid(
  top_combined,
  grid::textGrob("Patch size class", gp = grid::gpar(fontsize = 20)),
  ncol = 1, rel_heights = c(1, 0.05)
)

bottom_labeled <- cowplot::plot_grid(
  bottom_combined,
  grid::textGrob("Fire year", gp = grid::gpar(fontface = "bold", fontsize = 12)),
  ncol = 1, rel_heights = c(1, 0.05)
)

# Final stack
final_patch_plots <- cowplot::plot_grid(top_labeled, bottom_labeled, ncol = 1)
print(final_patch_plots)

# -------------------------------------------------------------------
# ANOVAs for patch metrics
# -------------------------------------------------------------------
# Core area ANOVA
aov_core <- aov(core_area_sum ~ rdnbr_class * n_bins, data = core_area_burned)
summary(aov_core)

# Perimeter-area ratio ANOVA
aov_paratio <- aov(perimeter_area_mean ~ rdnbr_class * n_bins, data = paratio_burned)
summary(aov_paratio)

# Contiguity ANOVA
aov_contiguity <- aov(contiguity_mean ~ rdnbr_class * n_bins, data = contiguity_burned)
summary(aov_contiguity)

# -------------------------------------------------------------------
# Assumptions checks
# -------------------------------------------------------------------
# Normality (example for core area)
shapiro.test(residuals(aov_core))
qqnorm(residuals(aov_core)); qqline(residuals(aov_core))

# Homogeneity (Levene's test)
car::leveneTest(core_area_sum ~ interaction(rdnbr_class, n_bins), data = core_area_burned)

# -------------------------------------------------------------------
# Tidy ANOVA outputs
# -------------------------------------------------------------------
core_tidy        <- broom::tidy(aov_core);        core_tidy$metric        <- "Core area"
paratio_tidy     <- broom::tidy(aov_paratio);     paratio_tidy$metric     <- "Perimeter-area ratio"
contiguity_tidy  <- broom::tidy(aov_contiguity);  contiguity_tidy$metric  <- "Contiguity"

anova_summary <- dplyr::bind_rows(core_tidy, paratio_tidy, contiguity_tidy)

print(anova_summary)

# Save boxplots (top row)
ggplot2::ggsave(
  filename = file.path(here::here(), "Figures/patches_boxes_new.png"),
  plot     = top_labeled,
  width    = 10, height = 7.5, units = "in", dpi = 300
)


```

#Table S5: Patch metrics summary

```{r}
# Bin sizes and compute area (ha)
patch_metrics_mega <- patch_metrics_mega %>%
  mutate(
    n_bins = cut(
      n,
      breaks = c(1, 100, 1000, 10000, 1500000),
      labels = c("small", "medium", "large", "very large"),
      include.lowest = TRUE
    ),
    area_ha = n * 0.09
  )

# Total area + patch counts by severity × size bin
patch_summary <- patch_metrics_mega %>%
  group_by(rdnbr_class, n_bins) %>%
  summarise(
    total_area  = sum(area_ha, na.rm = TRUE),
    patch_count = n(),
    .groups = "drop"
  )

# Mean / SD of patch area by severity × size bin
patch_stats <- patch_metrics_mega %>%
  group_by(rdnbr_class, n_bins) %>%
  summarise(
    mean_area = mean(area_ha, na.rm = TRUE),
    sd_area   = sd(area_ha, na.rm = TRUE),
    .groups   = "drop"
  )

# Patch metrics (means / SDs) by severity × size bin
patch_metrics_summary <- patch_metrics_mega %>%
  group_by(rdnbr_class, n_bins) %>%
  summarise(
    mean_contiguity      = mean(contiguity, na.rm = TRUE),
    sd_contiguity        = sd(contiguity, na.rm = TRUE),
    mean_perimeter_area  = mean(perimeter_area, na.rm = TRUE),
    sd_perimeter_area    = sd(perimeter_area, na.rm = TRUE),
    mean_core_area       = mean(core_area, na.rm = TRUE),
    sd_core_area         = sd(core_area, na.rm = TRUE),
    .groups              = "drop"
  ) %>%
  mutate(
    # keep formatting step (character output) as in your original
    mean_core_area = format(mean_core_area, scientific = FALSE),
    sd_core_area   = format(sd_core_area, scientific = FALSE)
  )

# Convenience subsets
patch_metrics_mega_vlarge     <- patch_metrics_mega %>% filter(n_bins == "very large")
patch_metrics_mega_vlarge_sev <- patch_metrics_mega_vlarge %>% filter(rdnbr_class == 3)
patch_metrics_mega_sev        <- patch_metrics_mega %>% filter(rdnbr_class == 3)


```

# Vegetation histograms



# All megafires

```{r}
# Colors for cover types
colors_all <- c('1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58', '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12', '8' = '#7b8d6a', '9' = '#e97451')

# Create empty dataframe

veg_sev_counts_master <- data.frame(
  rdnbr_class = integer(),
  RF_pre_veg = factor(),
  count=integer(),
  proportion = numeric(),
  OBJECTID = integer()
)

# -------------------------------------------------------------------------------------------------

for (num in megafire_IDs){
  
pixel_filename <- paste0(here(),"/Data/raster_df_mega_repeated_planting/",num,".csv")
pixel_df <- fread(pixel_filename, header=TRUE)  


# Calculate the proportions of RF_pre_veg within each rdnbr_class
pixel_df_summary <- pixel_df %>%
  group_by(rdnbr_class, RF_pre_veg) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count)) %>%
  ungroup()

pixel_df_summary$RF_pre_veg <- as.factor(pixel_df_summary$RF_pre_veg)  
pixel_df_summary$OBJECTID <-num

veg_sev_counts_master <- rbind(veg_sev_counts_master, pixel_df_summary)}


# Calculate the proportions of RF_pre_veg within each rdnbr_class
veg_sev_counts_summary <- veg_sev_counts_master %>%
  group_by(rdnbr_class, RF_pre_veg) %>%
  summarize(total_count = sum(count, na.rm = TRUE))%>%
  mutate(proportion = total_count / sum(total_count)) %>%
  ungroup()

veg_sev_counts_summary <- veg_sev_counts_summary %>%
  filter(rdnbr_class != -1)

# Create the stacked bar chart with custom colors, proportions
ggplot(veg_sev_counts_summary, aes(x = rdnbr_class, y = proportion, fill = RF_pre_veg)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = colors_all) +
  labs(title = "Stacked Bar Chart of RF_pre_veg Proportions by rdnbr_class",
       x = "RDNBR Class",
       y = "Proportion",
       fill = "RF_pre_veg") +
  theme_classic()

# Create the stacked bar chart with custom colors, absolute numbers
ggplot(veg_sev_counts_summary, aes(x = rdnbr_class, y = total_count, fill = RF_pre_veg, alpha=0.5)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = colors_all) +
  labs(title = "",
       x = "RDNBR Class",
       y = "Area",
       fill = "RF_pre_veg") +
  theme_bw()

# Only vegetation

veg_sev_counts_summary_vegOnly <- veg_sev_counts_summary %>%
  filter(RF_pre_veg %in% c("1", "4", "5","7","8","9"))

veg_sev_counts_summary_vegOnly$total_area_ha <- veg_sev_counts_summary_vegOnly$total_count*0.09

ggplot(veg_sev_counts_summary_vegOnly, aes(x = factor(rdnbr_class), y = total_area_ha, fill = RF_pre_veg)) +
  geom_bar(stat = "identity", position = "stack",alpha=1) +
scale_fill_manual(
    values = colors_all,
    name = "Pre-fire vegetation",
    labels = c("shrub","herbaceous","conifer", "sagebrush", "dense woodland", "oak woodland")
  ) +
  labs(
    title = "",
    x = "Burn severity (RdNBR)",
    y = "Area burned (ha)"
  ) +
  scale_x_discrete(
    labels = c('0' = 'unburned', '1' = 'low', '2' = 'moderate', '3' = 'high')
  ) +
  scale_y_continuous(labels = scales::comma) +  # Add thousands separators
  theme_classic() +
  theme(
    axis.title.x = element_text(margin = margin(t = 10),size=12, face='bold'),  # Add space on top of x-axis label
    axis.title.y = element_text(margin = margin(r = 10),size= 12, face='bold'),  # Add space to the right of y-axis label
    axis.text.x = element_text(angle = 0, hjust = 1,size=12),
    axis.text.y = element_text(size=12),# Angle x-axis labels to 45 degrees
    legend.position = c(0.01, 0.95),  # Move the legend inside the plot, top left
    legend.justification = c(0, 1),  # Adjust the alignment of the legend
    legend.title = element_text(size = 12), #, face = "bold"
    legend.text = element_text(size = 10),
    panel.grid = element_blank()  # Remove all gridlines
  )

```

#IGNORE Prop transitioned

```{r}
library(data.table)
library(fs)

# # Set path if using all fires, otherwise use below for only those with 20 years post-fire data
# csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"
# 
# # List all CSV files
# csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)

# Vector of fire IDs to include
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# Set path
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"

# List CSVs that match the fire IDs
csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]


# Initialize list to store results
summary_list <- list()

# Loop through each file
for (file in csv_files) {
  # Extract OBJECTID from filename (remove directory and extension)
  object_id <- path_ext_remove(path_file(file))
  
  # Read CSV with fread
  dt <- fread(file)
  
  # Filter and summarize
  dt_filtered <- dt[RF_pre_veg == 7, .N, by = .(rdnbr_class, transitioned)]
  dt_filtered[, OBJECTID := object_id]
  
  # Append to list
  summary_list[[length(summary_list) + 1]] <- dt_filtered
}

# Combine all summaries
final_summary <- rbindlist(summary_list)

# Optional: reshape to wide format
final_summary_wide <- dcast(
  final_summary,
  OBJECTID + rdnbr_class ~ transitioned,
  value.var = "N",
  fill = 0
)

# Rename columns for clarity
setnames(final_summary_wide, c("0", "1"), c("transitioned_0", "transitioned_1"))

# View result
print(final_summary_wide)


all_fires_transitioned <- final_summary_wide

all_fires_transitioned <- all_fires_transitioned %>%
  group_by(rdnbr_class) %>%
  summarise(count_trans = sum(transitioned_1), count_no_trans = sum(transitioned_0))




library(ggplot2)
library(tidyr)
library(dplyr)


# Reshape to long format
df_long <- all_fires_transitioned %>%
  pivot_longer(cols = c(count_trans, count_no_trans),
               names_to = "transition_status",
               values_to = "count") %>%
 mutate(transition_status = dplyr::recode(transition_status,
                                         "count_trans" = "Transitioned",
                                         "count_no_trans" = "Not Transitioned"))



# Calculate total and percentage per rdnbr_class
df_long <- df_long %>%
  group_by(rdnbr_class) %>%
  mutate(total = sum(count),
         percent = 100 * count / total) %>%
  ungroup()


# Prepare labels again (no changes here)
labels_df <- df_long %>%
  group_by(rdnbr_class) %>%
  summarise(
    total = sum(count),
    percent_trans = 100 * count[transition_status == "Transitioned"] / total
  ) %>%
  mutate(
    label = paste0(round(percent_trans, 1), "%")
  )

# Define color palette (unchanged)
custom_colors <- c("Not Transitioned" = "#0e4f12", "Transitioned" = "#e67e22")

# Reorder factor explicitly before plotting
df_long$transition_status <- factor(df_long$transition_status, 
                                    levels = c("Not Transitioned", "Transitioned"))

df_long <- df_long %>%
  arrange(rdnbr_class, transition_status)


ggplot(df_long, aes(x = factor(rdnbr_class), y = count, fill = transition_status)) +
  geom_bar(stat = "identity") +
  geom_text(data = labels_df,
            aes(x = factor(rdnbr_class), y = total, label = label),
            vjust = -0.5, size = 3.5, inherit.aes = FALSE) +
  scale_fill_manual(values = custom_colors) +
  scale_x_discrete(labels = c("0" = "Unchanged", "1" = "Low", "2" = "Moderate", "3" = "High")) +
  labs(
    x = "Burn severity class",
    y = "Pixel count",
    fill = "Transition status",
    title = "Transitioned vs. Non-Transitioned Pixels by Burn Severity"
  ) +
  theme_classic()


df_long <- df_long %>%
  mutate(
    transition_status = factor(transition_status, levels = c("Not Transitioned", "Transitioned")),
    area_ha = count * 0.00009
  )

df_stacked <- df_long %>%
  arrange(rdnbr_class, transition_status) %>%
  group_by(rdnbr_class) %>%
  mutate(
    ymin = cumsum(lag(area_ha, default = 0)),
    ymax = ymin + area_ha
  )

labels_df <- df_long %>%
  group_by(rdnbr_class) %>%
  summarise(
    total_area = sum(area_ha),
    percent_trans = 100 * area_ha[transition_status == "Transitioned"] / total_area
  ) %>%
  mutate(
    label = paste0(round(percent_trans, 1), "%")
  )


ggplot(df_stacked) +
  geom_rect(aes(xmin = as.numeric(factor(rdnbr_class)) - 0.4,
                xmax = as.numeric(factor(rdnbr_class)) + 0.4,
                ymin = ymin,
                ymax = ymax,
                fill = transition_status)) +
  geom_text(data = labels_df,
            aes(x = as.numeric(factor(rdnbr_class)), y = total_area, label = label),
            vjust = -0.5, size = 3.5, inherit.aes = FALSE) +
  scale_fill_manual(values = custom_colors) +
  scale_x_continuous(breaks = 1:4, labels = c("Unchanged", "Low", "Moderate", "High")) +
  labs(
    x = "Burn severity class",
    y = "Area (1,000 x ha)",
    fill = "Post-fire transition status",
    title = ""
  ) +
  theme_classic()


```

# IGNORE Prop returned

```{r}
# Prop returned

library(data.table)
library(fs)

# Vector of fire IDs to include
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# Set path
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"

# List CSVs that match the fire IDs
csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Initialize list to store summaries
summary_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  
  dt <- fread(file)
  
  # Only interested in conifer pre-fire pixels
  dt <- dt[RF_pre_veg == 7]
  
  dt <- dt[transitioned == 1]
  
  # Add recovery column
  dt[, returned_20yrs := fifelse(returned == 1 & yrs_to_return < 21, 1, 0)]
  
  # Summarize by rdnbr_class
  dt_summary <- dt[, .N, by = .(rdnbr_class, returned_20yrs)]
  dt_summary[, fire_id := fire_id]
  
  summary_list[[length(summary_list) + 1]] <- dt_summary
}

# Combine all
final_summary <- rbindlist(summary_list)

# Reshape wide
final_summary_wide <- dcast(
  final_summary,
  fire_id + rdnbr_class ~ returned_20yrs,
  value.var = "N",
  fill = 0
)

# Rename for clarity
setnames(final_summary_wide, old = c("0", "1"), new = c("not_returned", "returned"))

# Aggregate across fires
recovery_summary <- final_summary_wide %>%
  group_by(rdnbr_class) %>%
  summarise(
    count_returned = sum(returned),
    count_not_returned = sum(not_returned),
    .groups = "drop"
  )


# Prepare long format
df_long <- recovery_summary %>%
  pivot_longer(cols = c(count_returned, count_not_returned),
               names_to = "recovery_status",
               values_to = "count") %>%
 mutate(recovery_status = dplyr::recode(recovery_status,
                                "count_returned" = "Recovered ≤ 20 yrs",
                                "count_not_returned" = "Not Recovered"))


# Calculate area and percentages
df_long <- df_long %>%
  group_by(rdnbr_class) %>%
  mutate(
    total = sum(count),
    percent = 100 * count / total
  ) %>%
  ungroup()

df_long$area_ha <- df_long$count * 0.00009

# Calculate stacked bar coordinates
df_stacked <- df_long %>%
  arrange(rdnbr_class, recovery_status) %>%
  group_by(rdnbr_class) %>%
  mutate(
    ymin = cumsum(lag(area_ha, default = 0)),
    ymax = ymin + area_ha
  )

# Labels
labels_df <- df_long %>%
  group_by(rdnbr_class) %>%
  summarise(
    total_area = sum(area_ha),
    percent_recovered = 100 * area_ha[recovery_status == "Recovered ≤ 20 yrs"] / total_area
  ) %>%
  mutate(label = paste0(round(percent_recovered, 1), "%"))

# Plot
ggplot(df_stacked) +
  geom_rect(aes(xmin = as.numeric(factor(rdnbr_class)) - 0.4,
                xmax = as.numeric(factor(rdnbr_class)) + 0.4,
                ymin = ymin,
                ymax = ymax,
                fill = recovery_status)) +
  geom_text(data = labels_df,
            aes(x = as.numeric(factor(rdnbr_class)), y = total_area, label = label),
            vjust = -0.5, size = 3.5, inherit.aes = FALSE) +
  scale_fill_manual(values = c("Not Recovered" = "#d95f02", "Recovered ≤ 20 yrs" = "#1b9e77")) +
  scale_x_continuous(breaks = 1:4, labels = c("Unchanged", "Low", "Moderate", "High")) +
  labs(
    x = "Burn severity class",
    y = "Area (1,000 x ha)",
    fill = "Recovery within 20 years",
    title = ""
  ) +
  theme_classic()


```

# Fig. 4A

```{r}
# Load libraries
library(data.table)
library(fs)
library(dplyr)
library(tidyr)
library(ggplot2)

# Define fire IDs
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# Set path to CSVs
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store results
summary_list <- list()

# Loop over each fire
for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt <- fread(file)
  
  # Keep pre-fire conifer pixels
  dt <- dt[RF_pre_veg == 7]
  
  # Categorize recovery status
  dt[, recovery_status := fifelse(
    transitioned == 0, "Not Transitioned",
    fifelse(returned == 1 & yrs_to_return < 21, "Recovered ≤ 20 yrs", "Not Recovered")
  )]
  
  # Summarize by rdnbr_class
  dt_summary <- dt[, .N, by = .(rdnbr_class, recovery_status)]
  dt_summary[, fire_id := fire_id]
  
  summary_list[[length(summary_list) + 1]] <- dt_summary
}

# Combine all summaries
final_summary <- rbindlist(summary_list)

# Aggregate by rdnbr_class and category
recovery_summary <- final_summary %>%
  group_by(rdnbr_class, recovery_status) %>%
  summarise(count = sum(N), .groups = "drop")

# Convert to thousands of hectares (0.09 ha per 30m pixel)
df_long <- recovery_summary %>%
  group_by(rdnbr_class) %>%
  mutate(
    total = sum(count),
    percent = 100 * count / total,
    area_kha = count * 0.09 / 1000  # thousands of hectares
  ) %>%
  ungroup()

# Order levels for consistent stacking
df_long$recovery_status <- factor(df_long$recovery_status, 
                                  levels = c("Not Transitioned", "Not Recovered", "Recovered ≤ 20 yrs"))

# Arrange for stacking geometry
df_stacked <- df_long %>%
  arrange(rdnbr_class, recovery_status) %>%
  group_by(rdnbr_class) %>%
  mutate(
    ymin = cumsum(lag(area_kha, default = 0)),
    ymax = ymin + area_kha
  )

# Define custom colors
custom_colors <- c(
  "Not Transitioned" = "white",
  "Not Recovered" = alpha("darkgrey"),
  "Recovered ≤ 20 yrs" = alpha("darkgreen")
)

# Create the plot
plot_1 <- ggplot(df_stacked) +
  geom_rect(aes(
    xmin = as.numeric(factor(rdnbr_class)) - 0.4,
    xmax = as.numeric(factor(rdnbr_class)) + 0.4,
    ymin = ymin,
    ymax = ymax,
    fill = recovery_status),
    color = "black"
  ) +
  scale_fill_manual(values = custom_colors) +
  scale_x_continuous(
    breaks = 1:4,
    labels = c("unchanged", "low", "moderate", "high")
  ) +
  labs(
    x = "Burn severity class",
    y = "Area (1,000 ha)",
    fill = "Recovery category",
    title = ""
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.position = "none"
  )

# Show plot
print(plot_1)

```

# Fig. 4A Including plantings/ reburns

```{r}
# Load libraries
library(data.table)
library(dplyr)
library(ggplot2)
library(scales)

# Define fire IDs and file paths
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store summaries
summary_list <- list()

# Loop through each CSV
for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt <- fread(file)

  # Keep pre-fire conifer pixels
  dt <- dt[RF_pre_veg == 7]

  # Assign recovery categories with reburn priority
  dt[, recovery_status := fifelse(
    reburn_20yrs == 1, "reburned",
    fifelse(transitioned == 0, "Not Transitioned",
    fifelse(tree_planting == 1 & returned == 1, "planted_recovered",
    fifelse(tree_planting == 1 & returned == 0, "planted_nrecovered",
    fifelse(tree_planting == 0 & returned == 1 & yrs_to_return < 21, "Recovered ≤ 20 yrs", "Not Recovered"))))
  )]

  # Summarize by severity and recovery
  dt_summary <- dt[, .N, by = .(rdnbr_class, recovery_status)]
  dt_summary[, fire_id := fire_id]

  summary_list[[length(summary_list) + 1]] <- dt_summary
}

# Combine summaries
final_summary <- rbindlist(summary_list)

# Aggregate and compute area (in 1,000 ha)
recovery_summary <- final_summary %>%
  group_by(rdnbr_class, recovery_status) %>%
  summarise(count = sum(N), .groups = "drop")

df_long <- recovery_summary %>%
  group_by(rdnbr_class) %>%
  mutate(
    total = sum(count),
    percent = 100 * count / total,
    area_kha = count * 0.09 / 1000  # 30m pixels → 0.09 ha → /1000 = thousands of hectares
  ) %>%
  ungroup()

# Set factor levels
df_long$recovery_status <- factor(df_long$recovery_status,
  levels = c("Not Transitioned",
             "Recovered ≤ 20 yrs","planted_recovered","reburned",
             "Not Recovered","planted_nrecovered"))


# Stacking info
df_stacked <- df_long %>%
  arrange(rdnbr_class, recovery_status) %>%
  group_by(rdnbr_class) %>%
  mutate(
    ymin = cumsum(lag(area_kha, default = 0)),
    ymax = ymin + area_kha
  )

# Colors
custom_colors <- c(
  "Not Transitioned" = "white",
  "Not Recovered" = alpha("darkgrey"),
  "Recovered ≤ 20 yrs" = alpha("darkgreen"),
  "planted_nrecovered" = "#FF9999",
  "planted_recovered" = "#33CC33",
  "reburned" = "red"
)

# Labels
recovery_labels <- c(
  "Not Transitioned" = "Not transitioned",
  "Not Recovered" = "Unplanted, not recovered",
  "Recovered ≤ 20 yrs" = "Unplanted, recovered ≤ 20 yrs",
  "planted_nrecovered" = "Planted, not recovered",
  "planted_recovered" = "Planted, recovered",
  "reburned" = "Reburned"
)

# Plot
plot_1 <- ggplot(df_stacked) +
  geom_rect(aes(
    xmin = as.numeric(factor(rdnbr_class)) - 0.4,
    xmax = as.numeric(factor(rdnbr_class)) + 0.4,
    ymin = ymin,
    ymax = ymax,
    fill = recovery_status),
    color = "black"
  ) +
  scale_fill_manual(values = custom_colors, labels = recovery_labels) +
  scale_x_continuous(
    breaks = 1:4,
    labels = c("unchanged", "low", "moderate", "high")
  ) +
  labs(
    x = "Burn severity class",
    y = "Area (1,000 ha)",
    fill = "Recovery category",
    title = ""
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "none"
  )


```

# Fig. 4B Prop returned by patch size

```{r}
library(data.table)
library(dplyr)
library(ggplot2)

# Fire IDs and file paths
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"
csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store patch-level summaries
patch_prop_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  
  dt <- fread(file)
  
  # Only conifer and transitioned pixels
  dt <- dt[RF_pre_veg == 7 & transitioned == 1]
  if (nrow(dt) == 0) next
  
  # Recovery within 20 years
  dt[, returned_20yrs := as.integer(returned == 1 & yrs_to_return < 21)]
  
  # Summarise per patch
  patch_summary <- dt[, .(
    fire_id = fire_id,
    n_total = .N,
    n_returned = sum(returned_20yrs, na.rm = TRUE)
  ), by = id]
  
  patch_summary[, prop_returned := n_returned / n_total]
  
  # Assign patch size and class
  patch_summary[, patch_size_class := cut(
    n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very large"),
    include.lowest = TRUE
  )]
  
  patch_prop_list[[length(patch_prop_list) + 1]] <- patch_summary
}

# Combine all patches
patch_recovery_df <- rbindlist(patch_prop_list)

# Remove any NA patch size classes
patch_recovery_df <- patch_recovery_df[!is.na(patch_size_class)]

plot_2 <- ggplot(patch_recovery_df, aes(x = patch_size_class, y = prop_returned)) +
  geom_boxplot(fill = "grey", alpha = 0.4, outlier.size = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 0.8, color = "black") +
  labs(
    title = "",
    x = "Patch size class",
    y = "Proportion conifer recovered",
    caption = ""
  ) +
  theme_classic()+
  theme(
    axis.title = element_text( size = 14), #face = "bold",
    axis.text = element_text( size = 12),
    legend.position = "none"
  )



```

# Fig. 4B Including plantings/ reburns

```{r}
library(data.table)
library(dplyr)
library(ggplot2)

# Fire IDs and file paths
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- dir(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store patch-level summaries
patch_prop_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  
  dt <- fread(file)
  
  # Only conifer and transitioned pixels and not reburned or planted
  dt <- dt[RF_pre_veg == 7 & transitioned == 1 & reburn_20yrs == 0 & tree_planting == 0]
  if (nrow(dt) == 0) next
  
  # Recovery within 20 years
  dt[, returned_20yrs := as.integer(returned == 1 & yrs_to_return < 21)]
  
  # Summarise per patch
  patch_summary <- dt[, .(
    fire_id = fire_id,
    n_total = .N,
    n_returned = sum(returned_20yrs, na.rm = TRUE)
  ), by = id]
  
  patch_summary[, prop_returned := n_returned / n_total]
  
  # Assign patch size and class
  patch_summary[, patch_size_class := cut(
    n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very large"),
    include.lowest = TRUE
  )]
  
  patch_prop_list[[length(patch_prop_list) + 1]] <- patch_summary
}

# Combine all patches
patch_recovery_df <- rbindlist(patch_prop_list)

# Remove any NA patch size classes
patch_recovery_df <- patch_recovery_df[!is.na(patch_size_class)]

plot_2 <- ggplot(patch_recovery_df, aes(x = patch_size_class, y = prop_returned)) +
  geom_boxplot(fill = "white", alpha = 0.4, outlier.size = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 0.8, color = "black") +
  labs(
    title = "",
    x = "Patch size class",
    y = "Proportion conifer recovered",
    caption = ""
  ) +
  theme_classic()+
  theme(
    axis.title = element_text( size = 14), #face = "bold",
    axis.text = element_text( size = 12),
    legend.position = "none"
  )


```

```{r}
# Burn severity map

library(ggspatial)

fire_outline <- st_read("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Data/3310_SN_megafires_shp_individual/2012/39327.shp")

mega_rast_class <- rast("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Data/3310_SN_rdnbr_classified_individual_mega/39327.tif")

# Convert raster to data frame for ggplot
raster_df <- as.data.frame(mega_rast_class, xy = TRUE)


# Rename the raster value column (assuming single-layer raster)
colnames(raster_df)[3] <- "value"  # Renaming the value column
# Replace all -1 with 0 in the 'value' column of raster_df
raster_df <- raster_df %>%
  mutate(value = ifelse(value == -1, 0, value))

# Plot the map with shapefile and raster
map_rdnbr <- ggplot() +
  # Add shapefile (Layer 1)
  geom_sf(data = fire_outline, aes(fill = "Layer 1"), color = "black", alpha = 0.5) +  
  scale_fill_manual(
    values = c("Layer 1" = "lightgrey"),
    name = "",  # Legend title
    labels = c("KNP Complex fire")  # Custom labels
  ) +
  # Add raster layer
  geom_raster(data = raster_df, aes(x = x, y = y, fill = as.factor(value))) +
  scale_fill_manual(
    values = c("0" = "lightgrey", "1" = "#F6BE00", "2" = '#21918c', "3" = "#440154"),
    name = "Burn severity (RdNBR)",  # Custom legend title
    labels = c("unburned", "low", "moderate", "high")  # Custom labels
  ) + 
  theme_bw() +
  labs(
    title = "",
    subtitle = ""
  ) +
  theme(
    legend.position = c(0.15, 0.17),  # Position the legend
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )+
     annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(0.4, "in"), pad_y = unit(0.19, "in"),
                         style = north_arrow_fancy_orienteering) + 
  annotation_scale(location = "bl", width_hint = 0.5, style = "bar") 



# Map ohne alles

# Plot the map with shapefile and raster
map_rdnbr <- ggplot() +
  # Add shapefile (Layer 1)
  geom_sf(data = fire_outline, aes(fill = "Layer 1"), color = "black", alpha = 0.5) +  
  scale_fill_manual(
    values = c("Layer 1" = "lightgrey"),
    name = "",  # Legend title
    labels = c("KNP Complex fire")  # Custom labels
  ) +
  # Add raster layer
  geom_raster(data = raster_df, aes(x = x, y = y, fill = as.factor(value))) +
  scale_fill_manual(
    values = c("0" = "lightgrey", "1" = "#F6BE00", "2" = '#21918c', "3" = "#440154"),
    name = "Burn severity (RdNBR)",  # Custom legend title
    labels = c("unburned", "low", "moderate", "high")  # Custom labels
  ) + 
  theme_classic() +
  labs(
    title = "",
    subtitle = ""
  )



```

```{r}
#Pre-/post-fire vegetation maps

veg_rast_1pre <- rast("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Data/veg_annual_individual/22057/2006_22057.tif")
veg_rast_1post <- rast("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Data/veg_annual_individual/22057/2008_22057.tif")

# Convert raster to data frame for ggplot
raster_df <- as.data.frame(veg_rast_1pre, xy = TRUE)

# Rename the raster value column (assuming single-layer raster)
colnames(raster_df)[3] <- "value"  # Renaming the value column

ggplot() + 
  geom_sf(data = fire_outline, aes(fill = "Layer 1"), color = "black", alpha = 0.5) +  
  scale_fill_manual(
    values = c("Layer 1" = "lightgrey"),
    name = "",  # Legend title
    labels = c("KNP Complex fire")  # Custom labels
  ) +
  geom_raster(data = raster_df, aes(x = x, y = y, fill = as.factor(value))) + 
  scale_fill_manual(
    values = c('1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58', '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12', '8' = '#7b8d6a', '9' = '#e97451'),
    name = "Vegetation type",  # Custom legend title
    labels = c("Shrub", "Rock", "Water", "Open/grassland", "Herbaceous","Barren soil","Coniferous")  # Custom labels
  ) + 
  coord_sf() +  # Use coord_sf() for spatial data
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.title = element_blank(),  # Remove axis titles
    axis.text = element_blank(),   # Remove axis text
    axis.ticks = element_blank(),  # Remove axis ticks
    legend.position = "right",     # Place legend on the right
    legend.title = element_text(size = 12, face = "bold"),  # Customize legend title
    legend.text = element_text(size = 10),  # Customize legend text
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")  # Customize plot title
  ) +
  #ggtitle("KNP Complex Fire pre-fire vegetation types") +
  annotation_scale(location = "bl", width_hint = 0.5, style="bar")  # Add a scalebar at the bottom left




# Post-fire vegetation

# Convert raster to data frame for ggplot
raster_df <- as.data.frame(veg_rast_1post, xy = TRUE)

# Rename the raster value column (assuming single-layer raster)
colnames(raster_df)[3] <- "value"  # Renaming the value column


ggplot() + 
  geom_sf(data = fire_outline, aes(fill = "Layer 1"), color = "black", alpha = 0.5) +  
  scale_fill_manual(
    values = c("Layer 1" = "lightgrey"),
    name = "",  # Legend title
    labels = c("KNP Complex fire")  # Custom labels
  ) +
  geom_raster(data = raster_df, aes(x = x, y = y, fill = as.factor(value))) + 
  scale_fill_manual(
    values = c('1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58', '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12', '8' = '#7b8d6a', '9' = '#e97451'),
    name = "Vegetation type",  # Custom legend title
    labels = c("Shrub", "Rock", "Water", "Open/grassland", "Herbaceous","Barren soil","Coniferous")  # Custom labels
  ) + 
  coord_sf() +  # Use coord_sf() for spatial data
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),  # Remove gridlines
    axis.title = element_blank(),  # Remove axis titles
    axis.text = element_blank(),   # Remove axis text
    axis.ticks = element_blank(),  # Remove axis ticks
    legend.position = "right",     # Place legend on the right
    legend.title = element_text(size = 12, face = "bold"),  # Customize legend title
    legend.text = element_text(size = 10),  # Customize legend text
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")  # Customize plot title
  ) +
  #ggtitle("KNP Complex Fire pre-fire vegetation types") +
  annotation_scale(location = "bl", width_hint = 0.5, style="bar")  # Add a scalebar at the bottom left




```

# RdNBR histograms by fire

```{r}
library(raster)
library(rgdal)
library(ggplot2)
library(sp)
library(sf)
library(rgeos)
library(dplyr)
library(here)
library(stringr)
library(mgcv)
library(e1071)

# Histograms by veg type

pixel_df <- fread("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/718.csv", header=TRUE)  
 
# Convert burn severity class to factor
pixel_df$rdnbr_class <- as.factor(pixel_df$rdnbr_class) 

pixels_only_veg <- subset(pixel_df, RF_pre_veg %in% c("1", "4", "5","7","8","9")) 
pixels_only_veg$RF_pre_veg <- as.factor(pixels_only_veg$RF_pre_veg)

ggplot(pixels_only_veg, aes(x = rdnbr, fill = RF_pre_veg)) +
    geom_histogram(position = "identity", bins = 100, alpha = 0.4) +
    xlim(c(-1000, 2000)) +
    xlab("mean burn severity [rdnbr]") +
    scale_fill_manual(values = c('7' = '#0e4f12','1' = '#956733', '4' = '#d8bf58', '5' = '#6dcd2b','8' = '#7b8d6a', '9' = '#e97451')) +
    theme_classic() +
    theme(strip.background = element_rect(fill = "lightgrey", color = "grey"), strip.text = element_text(color = "black")) +
    theme(legend.position = "none")+
  facet_wrap(~RF_pre_veg)






# I would like to produce this graph for all megafires- I don't think this works yet, might only be reading/ adding the first fire/ file

# Create empty dataframe

veg_counts_master <- data.frame(
  rdnbr = integer(),
  RF_pre_veg = factor(),
  OBJECTID = integer()
)

# -------------------------------------------------------------------------------------------------

for (num in megafire_IDs){
  
pixel_filename <- paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/",num,".csv")
pixel_df <- fread(pixel_filename, header=TRUE)  


# Calculate the proportions of RF_pre_veg within each rdnbr_class
pixel_df_summary <- pixel_df %>%
  dplyr::select(rdnbr, RF_pre_veg)

pixel_df_summary$RF_pre_veg <- as.factor(pixel_df_summary$RF_pre_veg)  
pixel_df_summary$OBJECTID <-num

veg_counts_master <- rbind(veg_counts_master, pixel_df_summary)
print(nrow(veg_counts_master))}

veg_counts_master <- subset(veg_counts_master, RF_pre_veg %in% c("1", "4", "5","7","8","9")) 

# Define a named vector with your desired labels
veg_labels <- c(
  '1' = "Shrub",
  '4' = "Open woodland",
  '5' = "Herbaceous",
  '7' = "Conifer",
  '8' = "Sage brush",
  '9' = "Dense woodland"
)

ggplot(veg_counts_master, aes(x = rdnbr, fill = RF_pre_veg)) +
    geom_histogram(position = "identity", bins = 100, alpha = 0.9) +
    xlim(c(-1000, 2000)) +
    xlab("Mean burn severity [RdNBR]") +
    scale_fill_manual(values = c('7' = '#0e4f12','1' = '#956733', '4' = '#d8bf58', '5' = '#6dcd2b','8' = '#7b8d6a', '9' = '#e97451')) +
    theme_classic() +
    theme(strip.background = element_rect(fill = "lightgrey", color = "grey"), strip.text = element_text(color = "black")) +
    theme(legend.position = "none")+
  facet_wrap(~RF_pre_veg, labeller = labeller(RF_pre_veg = veg_labels))
 
```

## Recovery

```{r}
# For 20 year post-fire series: patch size vs. perc. recov after 20 years

patch_metrics_mega <- patch_metrics_mega %>%
  mutate(n_bins = cut(n,
                      breaks = c(1, 100, 1000,10000,1500000),
                      labels = c("small","medium","large","very_large"),
                      include.lowest = TRUE))

# Select patches with at least 20 years post-fire history
filtered_patch_metrics <- patch_metrics_mega %>%
  filter(OBJECTID %in% megafire_IDs)

filtered_patch_metrics$prop_con_trans <- 1- (filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7)
filtered_patch_metrics$prop_orig_con_1 <- filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_3 <- filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_5 <- filtered_patch_metrics$count_RF_post5_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_10 <- filtered_patch_metrics$count_RF_post10_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_15 <- filtered_patch_metrics$count_RF_post15_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_20 <- filtered_patch_metrics$count_RF_post20_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$abs_con_notRecov_20yrs <- filtered_patch_metrics$count_RF_pre_veg_7 - filtered_patch_metrics$count_RF_post20_veg_7


filtered_patch_metrics_HS <- filtered_patch_metrics %>%
  filter(rdnbr_class==3)

filtered_patch_metrics_5con <- filtered_patch_metrics %>%
  filter(count_RF_pre_veg_7 > 4)

filtered_patch_metrics_HS_5con <- filtered_patch_metrics_HS %>%
  filter(count_RF_pre_veg_7 > 4)

filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  filter(rdnbr_class >0)

# Combine n_bins and rdnbr_class into a single factor for the x-axis
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = interaction(n_bins, rdnbr_class, sep = "_"))


# Add spacing to the groups by adding empty levels
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = as.factor(n_bins_rdnbr)) %>%
  arrange(n_bins_rdnbr, rdnbr_class) %>%
  mutate(
    n_bins_rdnbr = factor(n_bins_rdnbr, levels = c(levels(n_bins_rdnbr), "")),  # Add an empty level for gaps
    rdnbr_class = as.factor(rdnbr_class)  # Ensure rdnbr_class is a factor
  )


# Summarize data for the marginal bar plot
marginal_data <- filtered_patch_metrics_5con %>%
  group_by(n_bins_rdnbr) %>%
  summarize(total_abs_con_notRecov = sum(abs_con_notRecov_20yrs, na.rm = TRUE))

marginal_data$total_abs_con_notRecov_ha <- marginal_data$total_abs_con_notRecov * 0.09

# Inspect unique values in n_bins_rdnbr
unique_bins <- unique(filtered_patch_metrics_5con$n_bins_rdnbr)

# Reorder levels dynamically based on observed values
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = factor(n_bins_rdnbr, levels = c("small_1","small_2","small_3", "medium_1","medium_2","medium_3", "large_1","large_2","large_3","very_large_1","very_large_2","very_large_3")))  # Replace with actual levels if different

# Update marginal_data with correct levels
marginal_data <- filtered_patch_metrics_5con %>%
  group_by(n_bins_rdnbr) %>%
  summarize(total_abs_con_notRecov = sum(abs_con_notRecov_20yrs, na.rm = TRUE)) %>%
  mutate(total_abs_con_notRecov_ha = total_abs_con_notRecov * 0.09)

# Recreate boxplot with adjusted order
boxplot <- ggplot(filtered_patch_metrics_5con, aes(x = n_bins_rdnbr, y = prop_orig_con_20, fill = rdnbr_class)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +  # Slightly transparent boxplot
  geom_jitter(aes(color = rdnbr_class), width = 0.2, alpha = 0.5, size = 1) +  # Scatter points with jitter
  stat_summary(
    fun = mean,  # Calculate the mean
    geom = "point",  # Add points
    shape = 18,  # Use a diamond shape for the mean
    size = 3,  # Adjust size of the mean point
    color = "red"  # Set color to red
  ) +
  labs(
    x = "n_bins and RDNBR Class",
    y = "Proportion of Original Conifer (20 Years)"
  ) +
  ylim(c(0, 1)) +
  scale_x_discrete(drop = FALSE) +  # Keep empty levels for gaps
  scale_fill_manual(values = c("#F6BE00", "#21918c", "#440154", "#F94144")) +  # Custom box colors
  scale_color_manual(values = c("lightgrey", "lightgrey", "lightgrey", "lightgrey")) +  # Custom scatter colors
  theme_classic() +
  theme(
    legend.position = "none",  # Remove legend
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Bar plot with adjusted order
barplot <- ggplot(marginal_data, aes(x = n_bins_rdnbr, y = -total_abs_con_notRecov_ha)) +
  geom_bar(stat = "identity", fill = "lightgrey", alpha = 0.8) +
  labs(
    x = "",
    y = "Total Abs Con Not Recovered (20 Years, ha)"
  ) +
  scale_x_discrete(drop = FALSE) +  # Keep empty levels for gaps
  theme_classic() +
  theme(
    axis.text.x = element_blank(),  # Remove overlapping labels
    axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.y = element_text(angle = 45, hjust = 1)
  )

# Add a title
title <- ggdraw() + 
  draw_label(
    "Boxplot with Custom Order and Marginal Bar Plot",
    fontface = 'bold',
    size = 14,
    hjust = 0.5
  )

# Combine the plots
combined_plot <- plot_grid(
  boxplot, barplot,
  ncol = 1,
  rel_heights = c(3, 1)  # Adjust the relative heights
)

# Add title to the combined plot
final_plot <- plot_grid(title, combined_plot, ncol = 1, rel_heights = c(0.1, 1))

# Print the combined plot
print(final_plot)


```

## Fig. 6 Recovery predictions

```{r}
### old model ###
####### For all megafires after 2002

# Load recovery predictions made with RF model

new_data_with_predictions <- fread("C:/Users/jscho/Documents/Megafires-1985-2023/all_predictions.csv")

new_data_with_predictions <- new_data_with_predictions %>%
  filter(fire_year > 2001)

# Group by fire_year and predicted_class, then summarize the counts
class_counts <- new_data_with_predictions %>%
  group_by(fire_year, predicted_class) %>%
  summarise(count = n(), .groups = "drop")

class_counts$area <- class_counts$count * 0.09


####### Read in earlier megafire files as well
# Directory containing the CSV files

directory <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"

# Vector of fire IDs
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# Get a list of all CSV files in the directory
csv_files <- list.files(directory, pattern = "\\.csv$", full.names = TRUE)

# Filter files that contain any of the megafire IDs in their filename
filtered_files <- csv_files[sapply(csv_files, function(file) {
  fire_id <- as.numeric(str_extract(basename(file), "\\d+")) # Extract numeric ID from filename
  fire_id %in% fire_IDs
})]

# Read and bind the filtered files into one dataframe
data_pre2002 <- bind_rows(lapply(filtered_files, fread))

# Filter data based on conditions
data_pre2002 <- data_pre2002 %>%
  filter(RF_pre_veg == 7) %>%
  filter(transitioned == 1) %>%
  filter(rdnbr_class == 3) %>%
  filter(tree_planting == 0) 

# Calculate additional columns
data_pre2002$years_since_fire <- data_pre2002$fire_year - data_pre2002$previous_fire_year
data_pre2002$returned[!is.na(data_pre2002$yrs_to_return) & data_pre2002$yrs_to_return > 20] <- 0

# Replace NA values in years_since_fire with 100
data_pre2002 <- data_pre2002 %>%
  mutate(years_since_fire = ifelse(is.na(years_since_fire), 100, years_since_fire))


# Group by fire_year and predicted_class, then summarize the counts
class_counts_predicted <- new_data_with_predictions %>%
  group_by(fire_year, predicted_class) %>%
  summarise(count = n(), .groups = "drop")

class_counts_predicted$type <- 'predicted'


class_counts_measured <- data_pre2002 %>%
  group_by(fire_year,returned)%>%
  summarise(count = n(), .groups = 'drop')

class_counts_measured$type <- 'measured'

class_counts_predicted <- setNames(class_counts_predicted, c("fire_year", "returned","count","type"))
class_counts_measured <- setNames(class_counts_measured, c("fire_year", "returned","count","type"))
class_counts_predicted$returned <- as.numeric(class_counts_predicted$returned)

class_counts <- rbind(class_counts_measured,class_counts_predicted)

class_counts$area <- class_counts$count * 0.09


bars_recovered <- ggplot(class_counts, aes(x = fire_year, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkgrey", "darkgreen")) +  # Manually set the colors
  scale_alpha_manual(values = c(1, 0.5)) +       # Full opacity for one, semi-transparent for another
  labs(title = "",
       x = "Fire year",
       y = "Area (ha)",
       fill = "Predicted class",
       alpha = "Type") +
  scale_y_continuous(labels = comma) +   # ← Add this line for thousand separators
  xlim(c(1987,2019))+
  ylim(c(0,25000))+
  theme_pubr()




# With cumulative area curves

library(dplyr)
library(ggplot2)
library(scales)
library(ggpubr)

# Step 1: Compute cumulative sums by returned class and fire year
cumulative_area <- class_counts %>%
  group_by(returned, fire_year) %>%
  summarise(area = sum(area), .groups = "drop") %>%
  arrange(returned, fire_year) %>%
  group_by(returned) %>%
  mutate(cum_area = cumsum(area))

# Step 2: Create the original bar chart
bars_recovered <- ggplot(class_counts, aes(x = fire_year, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  
  # Step 3: Add cumulative line
  geom_line(data = cumulative_area, 
            aes(x = fire_year, y = cum_area, color = factor(returned), group = returned),
            size = 1.2, inherit.aes = FALSE) +
  
  # Aesthetic settings
  scale_fill_manual(values = c("0" = "darkgrey", "1" = "darkgreen"),
                    labels = c("Not Recovered", "Recovered"),
                    name = "Returned") +
  scale_color_manual(values = c("0" = "darkgrey", "1" = "darkgreen"), guide = "none") +
  scale_alpha_manual(values = c("measured" = 1, "predicted" = 0.5)) +
  scale_y_continuous(labels = comma) +
  labs(
    x = "Fire Year",
    y = "Area (ha)",
    title = "",
    alpha = "Type",
    fill = "Returned"
  ) +
  theme_bw()


bars_recovered <- ggplot(class_counts, aes(x = fire_year, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +

  # Add cumulative line
  geom_line(data = cumulative_area, 
            aes(x = fire_year, y = cum_area, color = factor(returned), group = returned),
            size = 1.2, inherit.aes = FALSE) +

  # Manual color and alpha settings
  scale_fill_manual(values = c("0" = "darkgrey", "1" = "darkgreen"),
                    labels = c("Not Recovered", "Recovered"),
                    name = "Returned") +
  scale_color_manual(values = c("0" = "darkgrey", "1" = "darkgreen"), guide = "none") +
  scale_alpha_manual(values = c("measured" = 1, "predicted" = 0.5)) +

  # Scale y-axis by 1000 and remove scientific notation
  scale_y_continuous(
    labels = function(x) format(x / 1000, big.mark = ","),
    name = "Area (x1000 ha)"
  ) +

  # Axis titles and theme customization
  labs(
    x = "Fire year",
    title = "",
    alpha = "Type",
    fill = "Returned"
  ) +
  theme_bw() +  # start with minimal theme to remove gridlines
  theme(
    panel.grid = element_blank(),  # remove all gridlines
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16, face = "bold"),
    legend.position = "none"  # 👈 removes the legend
  )

# Step 4: Show the plot
print(bars_recovered)



  class_counts_prop <- class_counts %>%
    group_by(fire_year)%>%
    summarise(total_area = sum(area))

  merged_data <- class_counts %>%
    inner_join(class_counts_prop, by = "fire_year")

  merged_data_0 <- merged_data %>%
    filter(returned==0)

  merged_data_0$prop <- merged_data_0$area/merged_data_0$total_area


  ggplot(merged_data_0, aes(x = fire_year, y = prop)) +
    geom_point(size = 3, alpha = 0.8, color = "black") +  # Scatter plot
    geom_smooth(method = "lm", color = "black", se = TRUE) +  # Linear trendline
    labs(
      title = "Prop. severely burned conifer not recov/ pred. to recov w/in 20 yrs",
      x = "Fire year",
      y = "Proportion not recovered"
    ) +
    theme_classic() +  # Clean theme
    theme(
      text = element_text(size = 12),
      axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels if needed
    )


  # Fit the linear model
  linear_model <- lm(prop ~ fire_year, data = merged_data_0)

  # Summary of the model
  summary(linear_model)


  
# Plot absolute area not recovered/ predicted to not recover over time
  
  class_counts_noreturn <- class_counts %>%
    filter(returned == 0)
  
    ggplot(class_counts_noreturn, aes(x = fire_year, y = count)) +
    geom_point(size = 3, alpha = 0.8, color = "black") +  # Scatter plot
    geom_smooth(method = "lm", color = "black", se = TRUE) +  # Linear trendline
    labs(
      title = "",
      x = "Fire year",
      y = "Proportion not recovered"
    ) +
      ylim(c(0, 700000))+
    theme_classic() +  # Clean theme
    theme(
      text = element_text(size = 12),
      axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels if needed
    )
  
  # Fit the linear model
  linear_model <- lm(count ~ fire_year, data = class_counts_noreturn)

  # Summary of the model
  summary(linear_model)
  

class_counts_noreturn <- class_counts %>%
  filter(returned == 0)

ggplot(class_counts_noreturn, aes(x = fire_year, y = count)) +
  geom_point(size = 3, alpha = 0.8, color = "black") +  # Scatter plot
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", se = TRUE) +  # ← GAM here
  labs(
    title = "",
    x = "Fire year",
    y = "Area not recovered"
  ) +
  ylim(c(0, 150000)) +
  theme_classic() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_text(hjust = 1)
  )

gam_model <- gam(count ~ s(fire_year), data = class_counts_noreturn)
summary(gam_model)



#### Decadal sums of area at risk of transitioning:

class_counts_noreturn_dec1 <- class_counts_noreturn %>%
  filter(fire_year >= 1985, fire_year <= 1994)

class_counts_noreturn_dec2 <- class_counts_noreturn %>%
  filter(fire_year >= 1995, fire_year <= 2004)

class_counts_noreturn_dec3 <- class_counts_noreturn %>%
  filter(fire_year >= 2005, fire_year <= 2014)

class_counts_noreturn_dec4 <- class_counts_noreturn %>%
  filter(fire_year >= 2015, fire_year <= 2024)

sum_noreturn_dec1 <- sum(class_counts_noreturn_dec1$area)
#11025.63
sum_noreturn_dec2 <- sum(class_counts_noreturn_dec2$area)
#7710.48
sum_noreturn_dec3 <- sum(class_counts_noreturn_dec3$area)
#36284.13
sum_noreturn_dec4 <- sum(class_counts_noreturn_dec4$area)
#226533.96



# Pivot to a wider format

library(dplyr)
library(tidyr)

# Recode returned column for clearer labels
class_counts_wide <- class_counts %>%
  mutate(returned = ifelse(returned == 1, "recovered", "transitioned")) %>%
  dplyr::select(fire_year, type, returned, area) %>%
  pivot_wider(
    names_from = returned,
    values_from = area
  )

# View the resulting wide-format table
print(class_counts_wide)

fwrite(class_counts_wide, "C:/Users/jscho/Documents/Megafires-1985-2023-Data/counts_area_transitioned_annual_predicted.csv")



trans <- sum(class_counts_wide$transitioned, na.rm = TRUE)
rec <- sum(class_counts_wide$recovered, na.rm = TRUE)
total <- trans+rec

trans/total
```

# Fig 6 Including (excluding) reburns and plantings

```{r}
### Model considering reburns within 20 years ###

####### For all megafires after 2002

# Load recovery predictions made with RF model

new_data_with_predictions <- fread("C:/Users/jscho/Documents/Megafires-1985-2023/all_predictions_repeated_planting.csv")

new_data_with_predictions <- new_data_with_predictions %>%
  filter(fire_year > 2001)

# Group by fire_year and predicted_class, then summarize the counts
class_counts <- new_data_with_predictions %>%
  group_by(fire_year, predicted_class) %>%
  summarise(count = n(), .groups = "drop")

class_counts$area <- class_counts$count * 0.09


####### Read in earlier megafire files as well
# Directory containing the CSV files

directory <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"

# Vector of fire IDs
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# Get a list of all CSV files in the directory
csv_files <- list.files(directory, pattern = "\\.csv$", full.names = TRUE)

# Filter files that contain any of the megafire IDs in their filename
filtered_files <- csv_files[sapply(csv_files, function(file) {
  fire_id <- as.numeric(str_extract(basename(file), "\\d+")) # Extract numeric ID from filename
  fire_id %in% fire_IDs
})]

# Read and bind the filtered files into one dataframe
data_pre2002 <- bind_rows(lapply(filtered_files, fread))

# Filter data based on conditions
data_pre2002 <- data_pre2002 %>%
  filter(RF_pre_veg == 7) %>%
  filter(transitioned == 1) %>%
  filter(rdnbr_class == 3) %>%
  filter(tree_planting == 0) %>%
  filter(reburn_20yrs == 0)

# Calculate additional columns
data_pre2002$years_since_fire <- data_pre2002$fire_year - data_pre2002$previous_fire_year
data_pre2002$returned[!is.na(data_pre2002$yrs_to_return) & data_pre2002$yrs_to_return > 20] <- 0

# Replace NA values in years_since_fire with 100
data_pre2002 <- data_pre2002 %>%
  mutate(years_since_fire = ifelse(is.na(years_since_fire), 100, years_since_fire))


# Group by fire_year and predicted_class, then summarize the counts
class_counts_predicted <- new_data_with_predictions %>%
  group_by(fire_year, predicted_class) %>%
  summarise(count = n(), .groups = "drop")

class_counts_predicted$type <- 'predicted'


class_counts_measured <- data_pre2002 %>%
  group_by(fire_year,returned)%>%
  summarise(count = n(), .groups = 'drop')

class_counts_measured$type <- 'measured'

class_counts_predicted <- setNames(class_counts_predicted, c("fire_year", "returned","count","type"))
class_counts_measured <- setNames(class_counts_measured, c("fire_year", "returned","count","type"))
class_counts_predicted$returned <- as.numeric(class_counts_predicted$returned)

class_counts <- rbind(class_counts_measured,class_counts_predicted)

class_counts$area <- class_counts$count * 0.09


bars_recovered <- ggplot(class_counts, aes(x = fire_year, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkgrey", "darkgreen")) +  # Manually set the colors
  scale_alpha_manual(values = c(1, 0.5)) +       # Full opacity for one, semi-transparent for another
  labs(title = "",
       x = "Fire year",
       y = "Area (ha)",
       fill = "Predicted class",
       alpha = "Type") +
  scale_y_continuous(labels = comma) +   # ← Add this line for thousand separators
  xlim(c(1987,2019))+
  ylim(c(0,25000))+
  theme_pubr()




# With cumulative area curves

library(dplyr)
library(ggplot2)
library(scales)
library(ggpubr)

# Step 1: Compute cumulative sums by returned class and fire year
cumulative_area <- class_counts %>%
  group_by(returned, fire_year) %>%
  summarise(area = sum(area), .groups = "drop") %>%
  arrange(returned, fire_year) %>%
  group_by(returned) %>%
  mutate(cum_area = cumsum(area))

# Step 2: Create the original bar chart
bars_recovered <- ggplot(class_counts, aes(x = fire_year, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  
  # Step 3: Add cumulative line
  geom_line(data = cumulative_area, 
            aes(x = fire_year, y = cum_area, color = factor(returned), group = returned),
            size = 1.2, inherit.aes = FALSE) +
  
  # Aesthetic settings
  scale_fill_manual(values = c("0" = "darkgrey", "1" = "darkgreen"),
                    labels = c("Not Recovered", "Recovered"),
                    name = "Returned") +
  scale_color_manual(values = c("0" = "darkgrey", "1" = "darkgreen"), guide = "none") +
  scale_alpha_manual(values = c("measured" = 1, "predicted" = 0.5)) +
  scale_y_continuous(labels = comma) +
  labs(
    x = "Fire Year",
    y = "Area (ha)",
    title = "",
    alpha = "Type",
    fill = "Returned"
  ) +
  theme_bw()


bars_recovered <- ggplot(class_counts, aes(x = fire_year, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +

  # Add cumulative line
  geom_line(data = cumulative_area, 
            aes(x = fire_year, y = cum_area, color = factor(returned), group = returned),
            size = 1.2, inherit.aes = FALSE) +

  # Manual color and alpha settings
  scale_fill_manual(values = c("0" = "darkgrey", "1" = "darkgreen"),
                    labels = c("Not Recovered", "Recovered"),
                    name = "Returned") +
  scale_color_manual(values = c("0" = "darkgrey", "1" = "darkgreen"), guide = "none") +
  scale_alpha_manual(values = c("measured" = 1, "predicted" = 0.5)) +

  # Scale y-axis by 1000 and remove scientific notation
  scale_y_continuous(
    labels = function(x) format(x / 1000, big.mark = ","),
    name = "Area (x1000 ha)"
  ) +

  # Axis titles and theme customization
  labs(
    x = "Fire year",
    title = "",
    alpha = "Type",
    fill = "Returned"
  ) +
  theme_bw() +  # start with minimal theme to remove gridlines
  theme(
    panel.grid = element_blank(),  # remove all gridlines
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16, face = "bold"),
    legend.position = "none"  # 👈 removes the legend
  )

# Step 4: Show the plot
print(bars_recovered)



  class_counts_prop <- class_counts %>%
    group_by(fire_year)%>%
    summarise(total_area = sum(area))

  merged_data <- class_counts %>%
    inner_join(class_counts_prop, by = "fire_year")

  merged_data_0 <- merged_data %>%
    filter(returned==0)

  merged_data_0$prop <- merged_data_0$area/merged_data_0$total_area


  ggplot(merged_data_0, aes(x = fire_year, y = prop)) +
    geom_point(size = 3, alpha = 0.8, color = "black") +  # Scatter plot
    geom_smooth(method = "lm", color = "black", se = TRUE) +  # Linear trendline
    labs(
      title = "Prop. severely burned conifer not recov/ pred. to recov w/in 20 yrs",
      x = "Fire year",
      y = "Proportion not recovered"
    ) +
    theme_classic() +  # Clean theme
    theme(
      text = element_text(size = 12),
      axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels if needed
    )


  # Fit the linear model
  linear_model <- lm(prop ~ fire_year, data = merged_data_0)

  # Summary of the model
  summary(linear_model)


  
# Plot absolute area not recovered/ predicted to not recover over time
  
  class_counts_noreturn <- class_counts %>%
    filter(returned == 0)
  
    ggplot(class_counts_noreturn, aes(x = fire_year, y = count)) +
    geom_point(size = 3, alpha = 0.8, color = "black") +  # Scatter plot
    geom_smooth(method = "lm", color = "black", se = TRUE) +  # Linear trendline
    labs(
      title = "",
      x = "Fire year",
      y = "Proportion not recovered"
    ) +
      ylim(c(0, 700000))+
    theme_classic() +  # Clean theme
    theme(
      text = element_text(size = 12),
      axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels if needed
    )
  
  # Fit the linear model
  linear_model <- lm(count ~ fire_year, data = class_counts_noreturn)

  # Summary of the model
  summary(linear_model)
  

class_counts_noreturn <- class_counts %>%
  filter(returned == 0)

ggplot(class_counts_noreturn, aes(x = fire_year, y = count)) +
  geom_point(size = 3, alpha = 0.8, color = "black") +  # Scatter plot
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", se = TRUE) +  # ← GAM here
  labs(
    title = "",
    x = "Fire year",
    y = "Area not recovered"
  ) +
  ylim(c(0, 150000)) +
  theme_classic() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_text(hjust = 1)
  )

gam_model <- gam(count ~ s(fire_year), data = class_counts_noreturn)
summary(gam_model)



#### Decadal sums of area at risk of transitioning:

class_counts_noreturn_dec1 <- class_counts_noreturn %>%
  filter(fire_year >= 1985, fire_year <= 1994)

class_counts_noreturn_dec2 <- class_counts_noreturn %>%
  filter(fire_year >= 1995, fire_year <= 2004)

class_counts_noreturn_dec3 <- class_counts_noreturn %>%
  filter(fire_year >= 2005, fire_year <= 2014)

class_counts_noreturn_dec4 <- class_counts_noreturn %>%
  filter(fire_year >= 2015, fire_year <= 2024)

sum_noreturn_dec1 <- sum(class_counts_noreturn_dec1$area)
#7,108.11 old model: 11,025.63
sum_noreturn_dec2 <- sum(class_counts_noreturn_dec2$area)
#3,898.8 old model: 7,710.48
sum_noreturn_dec3 <- sum(class_counts_noreturn_dec3$area)
#46,692.9 old model: 36,284.13
sum_noreturn_dec4 <- sum(class_counts_noreturn_dec4$area)
#236,256.9 old model: 226,533.96



# Pivot to a wider format

library(dplyr)
library(tidyr)

# Recode returned column for clearer labels
class_counts_wide <- class_counts %>%
  mutate(returned = ifelse(returned == 1, "recovered", "transitioned")) %>%
  dplyr::select(fire_year, type, returned, area) %>%
  pivot_wider(
    names_from = returned,
    values_from = area
  )

# View the resulting wide-format table
print(class_counts_wide)

fwrite(class_counts_wide, "C:/Users/jscho/Documents/Megafires-1985-2023-Data/counts_area_transitioned_annual_predicted.csv")



trans <- sum(class_counts_wide$transitioned, na.rm = TRUE)
rec <- sum(class_counts_wide$recovered, na.rm = TRUE)
total <- trans+rec

trans/total
```

## Transitioned area in individual fires

```{r}
library(dplyr)
library(data.table)
library(stringr)
library(ggplot2)
library(scales)
library(ggpubr)

# Load recovery predictions (post-2001)
new_data_with_predictions <- fread("C:/Users/jscho/Documents/Megafires-1985-2023/all_predictions.csv") %>%
  filter(fire_year > 2001)

# Group by OBJECTID and predicted class
class_counts_predicted <- new_data_with_predictions %>%
  group_by(OBJECTID, predicted_class) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(type = "predicted")

# Load earlier megafire data
directory <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

csv_files <- list.files(directory, pattern = "\\.csv$", full.names = TRUE)
filtered_files <- csv_files[sapply(csv_files, function(file) {
  fire_id <- as.numeric(str_extract(basename(file), "\\d+"))
  fire_id %in% fire_IDs
})]

# Read and bind the raster-level fire data
data_pre2002 <- bind_rows(lapply(filtered_files, fread)) %>%
  filter(RF_pre_veg == 7, transitioned == 1, rdnbr_class == 3) %>%
  mutate(
    years_since_fire = ifelse(is.na(fire_year - previous_fire_year), 100, fire_year - previous_fire_year),
    returned = ifelse(!is.na(yrs_to_return) & yrs_to_return > 20, 0, returned)
  )

# Group by OBJECTID and returned for measured class
class_counts_measured <- data_pre2002 %>%
  group_by(OBJECTID, returned) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(type = "measured")

# Standardize and bind both datasets
class_counts_predicted <- setNames(class_counts_predicted, c("OBJECTID", "returned", "count", "type"))
class_counts_predicted$returned <- as.numeric(class_counts_predicted$returned)

class_counts <- bind_rows(class_counts_measured, class_counts_predicted) %>%
  mutate(area = count * 0.09)

# Load fire name data
# Assumes fire_metrics_mega has columns: OBJECTID and FIRE_NAME
fire_metrics_mega <- fread("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Dataframes/Megafire_paper/SN_megafires_1985_2023.csv", header=TRUE)

# Merge fire names
class_counts_named <- class_counts %>%
  left_join(fire_metrics_mega %>% dplyr::select(OBJECTID, FIRE_NAME, YEAR_), by = "OBJECTID") %>%
  filter(!is.na(FIRE_NAME))

class_counts_named <- class_counts_named %>%
  mutate(FIRE_NAME = factor(FIRE_NAME, levels = unique(FIRE_NAME[order(YEAR_)])))


# Final plot by FIRE_NAME
bars_by_fire_name <- ggplot(class_counts_named, aes(x = FIRE_NAME, y = area, fill = factor(returned), alpha = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkgrey", "darkgreen")) +
  scale_alpha_manual(values = c(1, 0.5)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "",
    x = "Fire name",
    y = "Area (ha)",
    fill = "Predicted class",
    alpha = "Type"
  ) +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Show plot
print(bars_by_fire_name)











# Scatterplot version of this graph

library(ggplot2)
library(scales)

# Compute cumulative area by returned class
cumulative_area <- class_counts_named %>%
  group_by(returned, YEAR_) %>%
  summarise(area = sum(area), .groups = "drop") %>%
  arrange(returned, YEAR_) %>%
  group_by(returned) %>%
  mutate(cum_area = cumsum(area))




scatter_by_fire_year <- ggplot(class_counts_named, aes(x = YEAR_, y = area)) +
  geom_point(aes(color = factor(returned), alpha = type),
             size = 2, position = position_jitter(width = 0.4, height = 0)) +
  geom_line(data = cumulative_area, aes(x = YEAR_, y = cum_area, color = factor(returned)),
            size = 1.2, inherit.aes = FALSE) +  # Add cumulative lines
  scale_color_manual(
    values = c("0" = "darkgrey", "1" = "darkgreen"),
    labels = c("Not Recovered", "Recovered"),
    name = "Returned"
  ) +
  scale_alpha_manual(
    values = c("measured" = 1, "predicted" = 0.5),
    name = "Type"
  ) +
  scale_y_continuous(labels = comma) +
  labs(
    x = "Fire Year",
    y = "Area (ha)",
    title = ""
  ) +
  theme_pubr() +
  theme(
    legend.position = "right",
    axis.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold")
  )



# Show the updated plot
print(scatter_by_fire_year)









```

## Fig. X: Patches over time

```{r}

patch_metrics_mega <- patch_metrics_mega %>%
  mutate(n_bins = cut(n,
                      breaks = c(1, 100, 1000,10000,1500000),
                      labels = c("small","medium","large","very_large"),
                      include.lowest = TRUE))

# Custom RdNBR color palette
rdnbr_colors <- c("0" = "lightgrey", "1" = "#F6BE00", "2" = "#21918c", "3" = "#440154")

# Count for very_large patches by fire_year and rdnbr_class
very_large_counts <- patch_metrics_mega %>%
  filter(n_bins == "very_large") %>%
  count(fire_year, rdnbr_class)

large_patches_timeseries <- ggplot(very_large_counts, aes(x = fire_year, y = n, color = factor(rdnbr_class))) +
  geom_point(size = 3, alpha = 0.9) +
  geom_smooth(method = "gam", se = TRUE) +
  scale_color_manual(values = rdnbr_colors) +
  labs(
    title = "",
    x = "Fire year",
    y = "Count",
    color = "RdNBR class"
  ) +
  theme_classic()+
  theme(legend.position = "none")

```

#Fig 2: Core area example map KNP

```{r}
# --------------------
# Load and Prepare Data
# --------------------

# Load RdNBR-classified raster (categorical burn severity: 1=low, 2=moderate, 3=severe)
r_path <- "C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Data/Megafire_paper/RdNBR_classified/718.tif"
r <- raster(r_path)

# Convert raster to terra object for manipulation and plotting
rdnbr_rast <- rast(r_path)
rdnbr_df <- as.data.frame(rdnbr_rast, xy = TRUE)
names(rdnbr_df)[3] <- "rdnbr_class"
rdnbr_df$rdnbr_class <- as.factor(rdnbr_df$rdnbr_class)

# Load fire perimeter shapefile
fire_outline <- st_read("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Data/Megafire_paper/mega_indiv_shp/2021/Mega_718.shp")

# Define custom RdNBR colors and labels
rdnbr_colors <- c("1" = "#F6BE00", "2" = "#21918c", "3" = "#440154")
rdnbr_labels <- c("1" = "low", "2" = "moderate", "3" = "severe")

# -----------------------------
# Identify Core and Non-Core Areas (Class 3 only)
# -----------------------------

# Define core edge depth and neighborhood window
edge_depth <- 5
window_size <- edge_depth * 2 + 1
window <- matrix(1, nrow = window_size, ncol = window_size)
max_neighbors <- sum(window)

# Binary masks
r_class3 <- r == 3                 # Mask of class 3 (severe)
r_non_na <- !is.na(r)              # Mask of valid cells

# Focal operations to count neighbors
class3_counts <- focal(r_class3, w = window, fun = sum, na.rm = FALSE, pad = TRUE)
valid_counts  <- focal(r_non_na, w = window, fun = sum, na.rm = FALSE, pad = TRUE)

# Core mask: class 3 pixels with full valid neighborhood of class 3\score_mask <- (class3_counts == max_neighbors) & (valid_counts == max_neighbors)

# Step 5: Core pixels = where class3_counts == max_neighbors AND valid_counts == max_neighbors
core_mask <- (class3_counts == max_neighbors) & (valid_counts == max_neighbors)

# Raster of core class 3 pixels
core_class3_raster <- mask(r, core_mask, maskvalue = FALSE)
core_binary <- calc(core_class3_raster, fun = function(x) ifelse(x == 3, 1, NA))

# Identify non-core class 3 pixels
non_core_class3 <- mask(r_class3, core_binary, maskvalue = 1)

# Combine core and non-core class 3 into one raster:
# 2 = core, 1 = non-core, NA = everything else
combined <- overlay(core_binary, non_core_class3, fun = function(core, noncore) {
  ifelse(!is.na(core) & core == 1, 2,
         ifelse(!is.na(noncore) & noncore == 1, 1, NA))
})

# Convert combined raster to dataframe for ggplot
core_df <- as.data.frame(combined, xy = TRUE)
colnames(core_df)[3] <- "burn_type"
core_df <- core_df %>%
  filter(!is.na(burn_type)) %>%
  mutate(burn_type = factor(burn_type, levels = c(1, 2), labels = c("Non-core", "Core")))

# --------------------
# Plotting
# --------------------

# (A) Map of RdNBR classes
map_plot_updated <- ggplot() +
  geom_tile(data = rdnbr_df, aes(x = x, y = y, fill = rdnbr_class)) +
  scale_fill_manual(values = rdnbr_colors, labels = rdnbr_labels, name = "RdNBR class") +
  geom_sf(data = fire_outline, fill = NA, color = "grey", size = 0.7) +
  coord_sf() +
  annotation_scale(location = "bl", width_hint = 0.5, text_cex = 0.8) +
  annotation_north_arrow(location = "bl", which_north = "true", pad_x = unit(0.3, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.position = "none"
  )

# (B) Core vs. Non-core Map
core_plot <- ggplot(core_df, aes(x = x, y = y, fill = burn_type)) +
  geom_tile() +
  scale_fill_manual(values = c("Non-core" = "#440154", "Core" = "orange")) +
  geom_sf(data = fire_outline, fill = NA, color = "grey", size = 0.7, inherit.aes = FALSE) +  # ← FIXED HERE
  coord_sf() +
  annotation_scale(location = "bl", width_hint = 0.5, text_cex = 0.8) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.3, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  labs(x = "Longitude", y = "Latitude", fill = "Burn Type") +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.position = "none"
  )

# --------------------
# Combine and Stack Maps with Bar Plot
# --------------------

# Combine maps (B and C)
maps_combined <- plot_grid(
  map_plot_updated,
  core_plot,
  labels = c("B", "C"),
  ncol = 2,
  align = "hv"
)

# Align cumulative bar plot (A) and stack all
cum_1_aligned <- cum_1 +
  theme(plot.margin = margin(t = 5, r = 5, b = 5, l = 40))  # aligns left axis

final_layout <- plot_grid(
  cum_1_aligned,
  maps_combined,
  ncol = 1,
  rel_heights = c(1, 1.5),
  labels = c("A", "")
)

# Display full figure
print(final_layout)
```

# Fig 4C: Including plantings/ reburns

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(stringr)

# Fire IDs
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# File paths
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Initialize list
patch_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt <- fread(file)

  # Filter: conifer, transitioned, not planted, not reburned
  dt <- dt[RF_pre_veg == 7 & transitioned == 1 & tree_planting == 0 & reburn_20yrs == 0]
  if (nrow(dt) == 0) next

  # Recovery flag
  dt[, returned_20yrs := as.integer(returned == 1 & yrs_to_return < 21)]

  # Summarize at patch (id) level
  patch_summary <- dt[, .(
    fire_id = fire_id,
    rdnbr_class = first(rdnbr_class),  # assuming all pixels in patch share class
    n_total = .N,
    n_returned = sum(returned_20yrs, na.rm = TRUE)
  ), by = id]

  # Patch size class
  patch_summary[, patch_size_class := cut(
    n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very_large"),
    include.lowest = TRUE
  )]

  patch_summary[, abs_not_recovered := n_total - n_returned]

  patch_list[[length(patch_list) + 1]] <- patch_summary
}

# Combine all
patch_data <- rbindlist(patch_list)
patch_data <- patch_data[!is.na(patch_size_class) & rdnbr_class > 0]

# Combine patch size + severity
patch_data <- patch_data %>%
  mutate(n_bins_rdnbr = interaction(patch_size_class, rdnbr_class, sep = "_"))

# Summarize for plot
marginal_data <- patch_data %>%
  group_by(n_bins_rdnbr) %>%
  summarize(total_abs_notRecov = sum(abs_not_recovered, na.rm = TRUE)) %>%
  mutate(
    total_abs_notRecov_ha = total_abs_notRecov * 0.09,
    total_abs_notRecov_thousand_ha = total_abs_notRecov_ha / 1000,
    severity = factor(str_extract(n_bins_rdnbr, "\\d$"), levels = c("1", "2", "3"))
  )

# Bar plot (plot_3)
plot_3 <- ggplot(marginal_data, aes(x = n_bins_rdnbr, y = total_abs_notRecov_thousand_ha, fill = severity)) +
  geom_bar(stat = "identity", alpha = 0.6) +
  scale_fill_manual(values = c("1" = "#F6BE00", "2" = "#21918c", "3" = "#440154")) +
  labs(
    x = "",
    y = "Area not recovered w/in 20 yrs (1,000 x ha)"
  ) +
  scale_y_continuous(labels = scales::comma_format()) +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.position = 'none'
  )

print(plot_3)

```

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(stringr)

# Fire IDs and CSV directory
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store all pixel data
pixel_data_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt_full <- fread(file)  # read full data

  # Get patch size (from all pixels)
  patch_sizes <- dt_full[, .(patch_n_total = .N, rdnbr_class = first(rdnbr_class)), by = id]

  # Merge patch size and severity into each pixel
  dt <- merge(dt_full, patch_sizes, by = "id", all.x = TRUE)

  # Keep only pre-fire conifer & transitioned
  dt <- dt[RF_pre_veg == 7 & transitioned == 1]

  if (nrow(dt) == 0) next

  # Classify patch size
  dt[, patch_size_class := cut(
    patch_n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very_large"),
    include.lowest = TRUE
  )]

  # Only non-recovered pixels within 20 years
  dt <- dt[returned == 0 | yrs_to_return >= 21 | is.na(returned)]

  # Assign pixel category
  dt[, pixel_status := fifelse(reburn_20yrs == 1, "reburned",
                        fifelse(tree_planting == 1, "planted", "none"))]

  dt[, fire_id := fire_id]

  pixel_data_list[[length(pixel_data_list) + 1]] <- dt
}

# Combine all pixel data
all_pixels <- rbindlist(pixel_data_list,fill = TRUE)

# Rename rdnbr_class.x or rdnbr_class.y if necessary
if ("rdnbr_class.x" %in% names(all_pixels)) {
  setnames(all_pixels, "rdnbr_class.x", "rdnbr_class")
}
if ("rdnbr_class.y" %in% names(all_pixels)) {
  all_pixels[, rdnbr_class.y := NULL]  # remove if present
}


# Remove NA patch sizes and invalid severity
all_pixels <- all_pixels[!is.na(patch_size_class) & rdnbr_class > 0]

# Summarize area by group
plot_data <- all_pixels %>%
  group_by(patch_size_class, rdnbr_class, pixel_status) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(area_kha = count * 0.09 / 1000)  # convert to 1,000 ha

# Order patch sizes
plot_data$patch_size_class <- factor(plot_data$patch_size_class,
                                     levels = c("small", "medium", "large", "very_large"))

# Order burn severity and pixel status
plot_data$rdnbr_class <- factor(plot_data$rdnbr_class, levels = c(1, 2, 3))
plot_data$pixel_status <- factor(plot_data$pixel_status, levels = c("none", "planted", "reburned"))

# Create plot
plot_stacked <- ggplot(plot_data, aes(x = interaction(patch_size_class, rdnbr_class),
                                      y = area_kha, fill = pixel_status)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  scale_fill_manual(values = c("none" = "#999999",  # gray
                               "planted" = "#33CC33",  # green
                               "reburned" = "#FF3333")) +  # red
  labs(
    x = "Patch size × Burn severity",
    y = "Not recovered conifer area (1,000 ha)",
    fill = "Pixel status"
  ) +
  
  scale_y_continuous(labels = scales::comma_format()) +
  scale_x_discrete(labels = function(x) {
    x <- gsub("_", " × ", x)
    gsub("very_large", "very large", x)
  }) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(plot_stacked)

```

```{r}
# Load libraries
library(data.table)
library(dplyr)
library(ggplot2)
library(stringr)

# Fire IDs and file paths
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store all pixel data
pixel_data_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt_full <- fread(file)  # Read full data

  # Step 1: Get patch sizes from full dataset
  patch_sizes <- dt_full[, .(patch_n_total = .N, rdnbr_class = first(rdnbr_class)), by = id]

  # Step 2: Filter to relevant pixels
  dt <- dt_full[RF_pre_veg == 7 & transitioned == 1]
  if (nrow(dt) == 0) next

  # Step 3: Flag recovery within 20 years
  dt[, returned_20yrs := as.integer(returned == 1 & yrs_to_return < 21)]
  dt[, abs_not_recovered := as.integer(returned == 0 | yrs_to_return >= 21 | is.na(returned))]

  # Step 4: Add pixel-level status
  dt[, pixel_status := fifelse(reburn_20yrs == 1, "reburned",
                        fifelse(tree_planting == 1, "planted", "none"))]

  # Step 5: Merge patch size data
  dt <- merge(dt, patch_sizes, by = "id", all.x = TRUE)

  # Step 6: Fix rdnbr_class column if renamed
  if ("rdnbr_class.x" %in% names(dt)) {
    dt[, rdnbr_class := rdnbr_class.x]
    dt[, rdnbr_class.x := NULL]
  }
  if ("rdnbr_class.y" %in% names(dt)) {
    dt[, rdnbr_class.y := NULL]
  }

  # Step 7: Classify patch size
  dt[, patch_size_class := cut(
    patch_n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very large"),
    include.lowest = TRUE
  )]

  # Add fire ID and store
  dt[, fire_id := fire_id]
  pixel_data_list[[length(pixel_data_list) + 1]] <- dt
}

# Combine all data
all_pixels <- rbindlist(pixel_data_list, fill = TRUE)

# Now safe to filter
all_pixels <- all_pixels[!is.na(patch_size_class) & !is.na(rdnbr_class) & rdnbr_class > 0]

# Filter to unplanted and unreburned only
plot_data <- all_pixels[pixel_status == "none"]

# Summarize area not recovered within 20 years
marginal_data <- plot_data %>%
  group_by(patch_size_class, rdnbr_class) %>%
  summarise(
    total_abs_notRecov = sum(abs_not_recovered, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    total_abs_notRecov_ha = total_abs_notRecov * 0.09,
    total_abs_notRecov_thousand_ha = total_abs_notRecov_ha / 1000
  )

# Plot
plot_none <- ggplot(marginal_data, aes(x = patch_size_class, y = total_abs_notRecov_thousand_ha, fill = factor(rdnbr_class))) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.5) +
  scale_fill_manual(
    values = c("1" = "#F6BE00", "2" = "#21918c", "3" = "#440154"),
    name = "Burn severity",
    labels = c("1" = "low", "2" = "moderate", "3" = "high")
  ) +
  labs(
    x = "Patch size class",
    y = "Area not recovered w/in 20 yrs (1,000 ha)",
    title = ""
  ) +
  scale_y_continuous(labels = scales::comma_format()) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "none"
  )

# Show plot
print(plot_none)
plot_3 <- plot_none

```

```{r}
# Load libraries
library(data.table)
library(dplyr)
library(ggplot2)
library(stringr)

# Define fire IDs and file paths
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]

# Store all pixel data
pixel_data_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt_full <- fread(file)  # Read full data

  # Step 1: Get patch sizes based on full dataset
  patch_sizes <- dt_full[, .(patch_n_total = .N, rdnbr_class = first(rdnbr_class)), by = id]

  # Step 2: Filter to relevant pixels
  dt <- dt_full[RF_pre_veg == 7 & transitioned == 1]

  if (nrow(dt) == 0) next

  # Step 3: Flag recovery within 20 years
  dt[, returned_20yrs := as.integer(returned == 1 & yrs_to_return < 21)]
  dt[, abs_not_recovered := as.integer(returned == 0 | yrs_to_return >= 21 | is.na(returned))]

  # Step 4: Add pixel-level status
  dt[, pixel_status := fifelse(reburn_20yrs == 1, "reburned",
                        fifelse(tree_planting == 1, "planted", "none"))]

  # Step 5: Merge patch size data
  dt <- merge(dt, patch_sizes, by = "id", all.x = TRUE)

  # Step 6: Classify patch size
  dt[, patch_size_class := cut(
    patch_n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very large"),
    include.lowest = TRUE
  )]

  # Save fire ID
  dt[, fire_id := fire_id]

  pixel_data_list[[length(pixel_data_list) + 1]] <- dt
}

# Combine all fires' pixel data
all_pixels <- rbindlist(pixel_data_list, fill = TRUE)

# Keep only valid patch sizes and burn severity
all_pixels <- all_pixels[!is.na(patch_size_class) & !is.na(rdnbr_class) & rdnbr_class > 0]

# Filter to unplanted and unreburned only
plot_data <- all_pixels[pixel_status == "none"]

# Summarize area not recovered within 20 years
marginal_data <- plot_data %>%
  group_by(patch_size_class, rdnbr_class) %>%
  summarise(
    total_abs_notRecov = sum(abs_not_recovered, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    total_abs_notRecov_ha = total_abs_notRecov * 0.09,
    total_abs_notRecov_thousand_ha = total_abs_notRecov_ha / 1000
  )

# Plot
plot_none <- ggplot(marginal_data, aes(x = patch_size_class, y = total_abs_notRecov_thousand_ha, fill = factor(rdnbr_class))) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  scale_fill_manual(
    values = c("1" = "#F6BE00", "2" = "#21918c", "3" = "#440154"),
    name = "Burn severity",
    labels = c("1" = "low", "2" = "moderate", "3" = "high")
  ) +
  labs(
    x = "Patch size class",
    y = "Area not recovered w/in 20 yrs (1,000 ha)",
    title = ""
  ) +
  scale_y_continuous(labels = scales::comma_format()) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = 'none'
  )

# Show plot
print(plot_none)

plot_3 <- plot_none

```

# Fig 4C: Transition box and bar

```{r}
# For 20 year post-fire series: patch size vs. perc. recov after 20 years

patch_metrics_mega <- patch_metrics_mega %>%
  mutate(n_bins = cut(n,
                      breaks = c(1, 100, 1000,10000,1500000),
                      labels = c("small","medium","large","very_large"),
                      include.lowest = TRUE))

# Select patches with at least 20 years post-fire history
filtered_patch_metrics <- patch_metrics_mega %>%
  filter(OBJECTID %in% megafire_IDs)

filtered_patch_metrics$prop_con_trans <- 1- (filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7)
filtered_patch_metrics$prop_orig_con_1 <- filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_3 <- filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_5 <- filtered_patch_metrics$count_RF_post5_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_10 <- filtered_patch_metrics$count_RF_post10_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_15 <- filtered_patch_metrics$count_RF_post15_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_20 <- filtered_patch_metrics$count_RF_post20_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$abs_con_notRecov_20yrs <- filtered_patch_metrics$count_RF_pre_veg_7 - filtered_patch_metrics$count_RF_post20_veg_7


filtered_patch_metrics_HS <- filtered_patch_metrics %>%
  filter(rdnbr_class==3)

filtered_patch_metrics_5con <- filtered_patch_metrics %>%
  filter(count_RF_pre_veg_7 > 4)

filtered_patch_metrics_HS_5con <- filtered_patch_metrics_HS %>%
  filter(count_RF_pre_veg_7 > 4)

filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  filter(rdnbr_class >0)

# Combine n_bins and rdnbr_class into a single factor for the x-axis
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = interaction(n_bins, rdnbr_class, sep = "_"))


# Add spacing to the groups by adding empty levels
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = as.factor(n_bins_rdnbr)) %>%
  arrange(n_bins_rdnbr, rdnbr_class) %>%
  mutate(
    n_bins_rdnbr = factor(n_bins_rdnbr, levels = c(levels(n_bins_rdnbr), "")),  # Add an empty level for gaps
    rdnbr_class = as.factor(rdnbr_class)  # Ensure rdnbr_class is a factor
  )

# Summarize data for the marginal bar plot
marginal_data <- filtered_patch_metrics_5con %>%
  group_by(n_bins_rdnbr) %>%
  summarize(total_abs_con_notRecov = sum(abs_con_notRecov_20yrs, na.rm = TRUE))

marginal_data$total_abs_con_notRecov_ha <- marginal_data$total_abs_con_notRecov * 0.09

# Inspect unique values in n_bins_rdnbr
unique_bins <- unique(filtered_patch_metrics_5con$n_bins_rdnbr)

# Reorder levels dynamically based on observed values
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = factor(n_bins_rdnbr, levels = c("small_1","small_2","small_3", "medium_1","medium_2","medium_3", "large_1","large_2","large_3","very_large_1","very_large_2","very_large_3")))  # Replace with actual levels if different

# Update marginal_data with correct levels
marginal_data <- filtered_patch_metrics_5con %>%
  group_by(n_bins_rdnbr) %>%
  summarize(total_abs_con_notRecov = sum(abs_con_notRecov_20yrs, na.rm = TRUE)) %>%
  mutate(total_abs_con_notRecov_ha = total_abs_con_notRecov * 0.09)

# Recreate boxplot with adjusted order
boxplot <- ggplot(filtered_patch_metrics_5con, aes(x = n_bins_rdnbr, y = prop_orig_con_20, fill = rdnbr_class)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +  # Slightly transparent boxplot
  geom_jitter(aes(color = rdnbr_class), width = 0.2, alpha = 0.5, size = 1) +  # Scatter points with jitter
  stat_summary(
    fun = mean,  # Calculate the mean
    geom = "point",  # Add points
    shape = 18,  # Use a diamond shape for the mean
    size = 3,  # Adjust size of the mean point
    color = "red"  # Set color to red
  ) +
  labs(
    x = "n_bins and RDNBR Class",
    y = "Proportion of Original Conifer (20 Years)"
  ) +
  ylim(c(0, 1)) +
  scale_x_discrete(drop = FALSE) +  # Keep empty levels for gaps
  scale_fill_manual(values = c("#F6BE00", "#21918c", "#440154", "#F94144")) +  # Custom box colors
  scale_color_manual(values = c("lightgrey", "lightgrey", "lightgrey", "lightgrey")) +  # Custom scatter colors
  theme_classic() +
  theme(
    legend.position = "none",  # Remove legend
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


marginal_data <- marginal_data %>%
  mutate(
    severity = factor(
      stringr::str_extract(n_bins_rdnbr, "\\d$"),  # extract last digit
      levels = c("1", "2", "3")
    )
  )

barplot <- ggplot(marginal_data, aes(x = n_bins_rdnbr, y = -total_abs_con_notRecov_ha, fill = severity)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  scale_fill_manual(values = c("1" = "#F6BE00", "2" = "#21918c", "3" = "#440154")) +
  labs(
    x = "",
    y = "Total Abs Con Not Recovered (20 Years, ha)"
  ) +
  scale_x_discrete(drop = FALSE) +  # Keep empty levels for gaps
  theme_classic() +
  theme(
    axis.text.x = element_blank(),       # Remove overlapping labels
    axis.title.x = element_blank(),      # Remove x-axis title
    axis.text.y = element_text(angle = 45, hjust = 1),
    legend.position = 'none'
  )

print(barplot)



# Combine the plots
combined_plot <- plot_grid(
  boxplot, barplot,
  ncol = 1,
  rel_heights = c(3, 1)  # Adjust the relative heights
)

combined_plot <- plot_grid(
  boxplot, barplot,
  ncol = 1,
  rel_heights = c(3, 1),
  align = "v",       # vertically align the plots
  axis = "lr"        # align left and right axes
)

# Add title to the combined plot
final_plot <- plot_grid(combined_plot, ncol = 1, rel_heights = c(0.1, 1))



# Print the combined plot
print(final_plot)







# ANOVA

# Run one-way ANOVA- This tests if mean prop_orig_con_20 differs by group (i.e., small_1, medium_3, etc.).
anova_result <- aov(prop_orig_con_20 ~ n_bins_rdnbr, data = filtered_patch_metrics_5con)
summary(anova_result)


# Two-way ANOVA with interaction- This will show Main effects of patch size (n_bins), Main effects of burn severity (rdnbr_class), Interaction effects between the two

anova_interaction <- aov(prop_orig_con_20 ~ n_bins * rdnbr_class, data = filtered_patch_metrics_5con)
summary(anova_interaction)

# Normality of residuals
plot(anova_result, which = 2)  # Q-Q plot

# Homogeneity of variance
plot(anova_result, which = 1)  # Residuals vs fitted

# Levene's Test (optional, from car package)
library(car)
leveneTest(prop_orig_con_20 ~ n_bins_rdnbr, data = filtered_patch_metrics_5con)

TukeyHSD(anova_result)

TukeyHSD(anova_interaction)








# Only the barplot, right way up

marginal_data$total_abs_con_notRecov_thousand_ha <- marginal_data$total_abs_con_notRecov_ha/ 1000

plot_3 <- ggplot(marginal_data, aes(x = n_bins_rdnbr, y = total_abs_con_notRecov_thousand_ha, fill = severity)) +
  geom_bar(stat = "identity", alpha = 0.3) +
  scale_fill_manual(values = c("1" = "#F6BE00", "2" = "#21918c", "3" = "#440154")) +
  labs(
    x = "",
    y = "Area not recovered w/in 20 yrs (1,000 x ha)"
  ) +
  scale_y_continuous(labels = scales::comma_format())+
  scale_x_discrete(drop = FALSE) +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    legend.position = 'none',
    axis.title = element_text( size = 14), #face = "bold",
    axis.text = element_text( size = 12),
  )

print(plot_3)
```

# New figure 4

```{r}
library(cowplot)

# Combine plot_1 and plot_2 in top row
top_row <- plot_grid(plot_1, plot_2, labels = c("A", "B"), label_size = 14)

# Combine top row with plot_3 below
final_plot <- plot_grid(top_row, plot_3, ncol = 1, labels = c("", "C"), label_size = 14, rel_heights = c(1, 1))


library(cowplot)
# Combine plot_1 and plot_2, aligned horizontally
top_row <- plot_grid(
  plot_1, plot_2,
  labels = c("A", "B"),
  label_size = 14,
  align = "h",        # align horizontally
  axis = "b"          # align bottom axis lines
)

# Add plot_3 below
final_plot <- plot_grid(
  top_row, plot_3,
  ncol = 1,
  labels = c("", "C"),
  label_size = 14,
  rel_heights = c(1, 1)
)


```

# Mapping individual patches

```{r}

library(biscale)
library(scales)
library(ggspatial)

fire_id <- 10495
patch_id <- 3538
df_fire <- fread(paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/",fire_id,".csv"))

#find some large patches:

id_freq_rdnbr <- df_fire %>%
  group_by(id) %>%
  summarise(
    frequency = n(),
    rdnbr_class = first(rdnbr_class)  # or use `unique(rdnbr_class)` if paranoid about duplicates
  ) %>%
  arrange(desc(frequency))  # Optional: sort by frequency

print(id_freq_rdnbr)

# largest high severity patch is id == 3877

df_patch <- df_fire %>%
  filter(id == patch_id)


# Pre-fire vegetation types within patch

veg_freq <- df_patch %>%
  group_by(RF_pre_veg) %>%
  summarise(
    frequency = n() 
  ) %>%
  arrange(desc(frequency))  # Optional: sort by frequency

print(veg_freq)


### Plot bivariate map returned within 20 yrs vs. distance from unburned greater/ less than 150m

# Define fire and patch IDs
fire_id <- 10094
patch_id <- 869

# Load data dynamically based on fire ID
file_path <- paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/", fire_id, ".csv")
df_fire <- fread(file_path)

# Filter for specific patch and conifer vegetation
df_patch <- df_fire %>%
  filter(id == patch_id, RF_pre_veg == 7)

# Select necessary variables and remove NAs
df_map <- df_patch %>%
  select(x, y, returned, dist_unburned) %>%
  filter(!is.na(returned), !is.na(dist_unburned))

# Create binned variables for bivariate mapping
df_map <- df_map %>%
  mutate(
    dist_bin = ifelse(dist_unburned < 150, 1, 2),
    returned_bin = ifelse(returned == 0, 1, 2),  # 1 = not returned, 2 = returned
    bi_class = paste0(dist_bin, "-", returned_bin)
  )

# Create bivariate map
map_plot <- ggplot(df_map, aes(x = x, y = y, fill = bi_class)) +
  geom_raster(color = "white", size = 0.05) +
  bi_scale_fill(pal = "BlueGold", dim = 2, na.value = "darkgrey") +
  coord_equal() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  labs(x = "Longitude", y = "Latitude") +
  bi_theme() +
  theme(
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(face = "bold", size = 8),
    legend.position = "none"
  )

# Create legend
legend <- bi_legend(
  pal = "BlueGold",
  dim = 2,
  xlab = "Distance from unburned",
  ylab = "Forest Recovery",
  size = 8,
  arrow = TRUE
)

# Combine map and legend
final_map <- plot_grid(
  map_plot,
  legend,
  rel_widths = c(0.88, 0.12)
)

# Show it
print(final_map)



```

```{r}
# Find patches in fires before 2002 with rdnbr_class == 3, and more than 500 pixels RF_pre_veg == 7

library(data.table)
library(dplyr)

fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# List to store patch summaries
patch_summary_list <- list()

for (fire_id in fire_IDs) {
  
  # Load the CSV for the fire
  file_path <- paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/", fire_id, ".csv")
  if (!file.exists(file_path)) {
    warning(paste("File missing for fire", fire_id))
    next
  }
  
  df_fire <- fread(file_path)
  
  # Ensure needed columns exist
  if (!all(c("id", "rdnbr_class", "RF_pre_veg") %in% names(df_fire))) {
    warning(paste("Required columns missing in fire", fire_id))
    next
  }
  
  # Group by patch ID and summarise
  id_freq_rdnbr <- df_fire %>%
    group_by(id) %>%
    summarise(
      fire_id = fire_id,
      frequency = n(),
      rdnbr_class = first(rdnbr_class),
      pre_veg_7_count = sum(RF_pre_veg == 7, na.rm = TRUE),
      .groups = "drop"
    )
  
  # Filter to patches meeting criteria:
  selected_patches <- id_freq_rdnbr %>%
    filter(rdnbr_class == 3 & (frequency > 1000 | pre_veg_7_count > 500))
  
  # Store in list
  patch_summary_list[[as.character(fire_id)]] <- selected_patches
}

# Combine all into one dataframe
patch_summary_df <- bind_rows(patch_summary_list)

# View or export
print(patch_summary_df)
# write.csv(patch_summary_df, "patch_summary_selected.csv", row.names = FALSE)

```

```{r}
# Map the just found patches

library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(fs)

# Set path to fire CSVs
data_path <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"

# Color palette
veg_colors <- c('1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58', '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12', '8' = '#7b8d6a', '9' = '#e97451')

# Loop over each patch
for (i in 1:nrow(patch_summary_df)) {
  
  fire_id <- patch_summary_df$fire_id[i]
  patch_id <- patch_summary_df$id[i]
  
  # Load the fire raster dataframe
  df_file <- file.path(data_path, paste0(fire_id, ".csv"))
  if (!file_exists(df_file)) next
  
  df_fire <- fread(df_file)
  
  # Subset to the patch
  df_patch <- df_fire %>% filter(id == patch_id)
  
  # Extract all RF_postX_veg and RF_pre_veg columns
  veg_df <- df_patch %>%
    select(starts_with("RF_post"), RF_pre_veg)
  
  # Pivot to long format
  veg_long <- veg_df %>%
    pivot_longer(cols = everything(), names_to = "year_label", values_to = "veg_class") %>%
    mutate(
      rel_year = ifelse(year_label == "RF_pre_veg", 0,
                        as.integer(gsub("RF_post(\\d+)_veg", "\\1", year_label))),
      veg_class = as.character(veg_class)
    ) %>%
    group_by(rel_year, veg_class) %>%
    summarise(count = n(), .groups = "drop")
  
veg_long <- veg_long %>%
  mutate(
    veg_class = factor(veg_class, levels = c(as.character(sort(unique(veg_class[veg_class != "7"]))), "7"))

  ) %>%
  arrange(rel_year, veg_class)  # 👈 stacking order now guaranteed

used_levels <- levels(veg_long$veg_class)
veg_colors_ordered <- veg_colors[used_levels]

p <- ggplot(veg_long, aes(x = rel_year, y = count, fill = veg_class)) +
  geom_bar(stat = "identity", position = "stack", width = 0.7) +
  scale_fill_manual(values = veg_colors_ordered, drop = FALSE, na.translate = FALSE) +
  labs(
    title = paste("Vegetation Change - Fire", fire_id, "Patch", patch_id),
    x = "Years Since Fire (0 = Pre-fire)",
    y = "Pixel Count",
    fill = "Vegetation Class"
  ) +
  theme_classic(base_size = 12)


  
  # Save
  out_file <- paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/patch_veg/veg_trajectory_fire", fire_id, "_patch", patch_id, ".png")
  ggsave(out_file, plot = p, width = 8, height = 5, dpi = 300)
}

```

```{r}
# Actually maps

library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(fs)

# Set paths and color map
data_path <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"

veg_colors <- c('1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58', '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12', '8' = '#7b8d6a', '9' = '#e97451')

# Loop through each patch
for (i in 1:nrow(patch_summary_df)) {
  
  fire_id <- patch_summary_df$fire_id[i]
  patch_id <- patch_summary_df$id[i]
  
  # Load the fire CSV
  df_file <- file.path(data_path, paste0(fire_id, ".csv"))
  if (!file_exists(df_file)) next
  
  df_fire <- fread(df_file)
  
  # Subset to patch
  df_patch <- df_fire %>% filter(id == patch_id)
  
  # Pivot all RF_postX_veg and RF_pre_veg to long format
  veg_spatial <- df_patch %>%
    select(x, y, starts_with("RF_post"), RF_pre_veg) %>%
    pivot_longer(cols = starts_with("RF_"), names_to = "year_label", values_to = "veg_class") %>%
    mutate(
      rel_year = ifelse(year_label == "RF_pre_veg", 0,
                        as.integer(gsub("RF_post(\\d+)_veg", "\\1", year_label))),
      veg_class = as.character(veg_class)
    ) %>%
    filter(!is.na(veg_class))
  
  # Plot faceted tile map
  p <- ggplot(veg_spatial, aes(x = x, y = y, fill = veg_class)) +
    geom_tile() +
    scale_fill_manual(values = veg_colors, na.translate = FALSE) +
    coord_equal() +
    facet_wrap(~ rel_year) +
    labs(
      title = paste("Vegetation Structure Over Time\nFire", fire_id, "Patch", patch_id),
      fill = "Veg Class"
    ) +
    theme_bw() +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      strip.text = element_text(size = 10, face = "bold")
    )
  
  # Save map
  out_file <- paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/patch_veg_map/veg_patch_map_fire", fire_id, "_patch", patch_id, ".png")
  ggsave(out_file, plot = p, width = 10, height = 6, dpi = 300)
}

```

# Specific patch recovery

```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(fs)
library(grid)

# Fire/Patch selection
# selected_patches <- data.frame(
#   fire_id = c(7514, 7242, 7514),
#   patch_id = c(3512, 3158, 3902)
# )

selected_patches <- data.frame(
  fire_id = c(10273, 10273, 10498),
  patch_id = c(3647, 3692, 3279)
)

# Just before the join
selected_patches$fire_id <- as.integer(selected_patches$fire_id)
fire_metrics_mega$OBJECTID <- as.integer(fire_metrics_mega$OBJECTID)

selected_patches <- selected_patches %>%
  left_join(fire_metrics_mega, by = c("fire_id" = "OBJECTID")) %>%
  mutate(FIRE_NAME = ifelse(fire_id == 10273, "CLARK", FIRE_NAME)) %>%
  mutate(title = paste0(FIRE_NAME, " (", YEAR_, "), Patch ", patch_id))



data_path <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"

# Color and label definitions
veg_colors <- c(
  '1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58',
  '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12',
  '8' = '#7b8d6a', '9' = '#e97451'
)

veg_labels <- c(
  '1' = 'shrub', '4' = 'open woodland', '5' = 'herbaceous',
  '7' = 'conifer', '8' = 'sagebrush', '9' = 'dense woodland'
)

# Store individual plots
plot_list <- list()

# Loop
for (i in 1:nrow(selected_patches)) {
  fire_id <- selected_patches$fire_id[i]
  patch_id <- selected_patches$patch_id[i]
  title_text <- selected_patches$title[i]

  df_file <- file.path(data_path, paste0(fire_id, ".csv"))
  if (!file_exists(df_file)) next

  df_fire <- fread(df_file)
  df_patch <- df_fire %>% filter(id == patch_id)

  veg_df <- df_patch %>% select(starts_with("RF_post"), RF_pre_veg)

  veg_long <- veg_df %>%
    pivot_longer(cols = everything(), names_to = "year_label", values_to = "veg_class") %>%
    mutate(
      rel_year = ifelse(year_label == "RF_pre_veg", 0,
                        as.integer(gsub("RF_post(\\d+)_veg", "\\1", year_label))),
      veg_class = as.character(veg_class)
    ) %>%
    group_by(rel_year, veg_class) %>%
    summarise(count = n(), .groups = "drop") %>%
    mutate(
      veg_class = factor(veg_class, levels = c(sort(unique(veg_class[veg_class != "7"])), "7"))
    ) %>%
    arrange(rel_year, veg_class)

  used_levels <- levels(veg_long$veg_class)
  veg_colors_ordered <- veg_colors[used_levels]
  veg_labels_ordered <- veg_labels[used_levels]
  

  p <- ggplot(veg_long, aes(x = rel_year, y = count, fill = veg_class)) +
    geom_bar(stat = "identity", position = "stack", width = 0.7) +
    scale_fill_manual(values = veg_colors_ordered,
                      labels = veg_labels_ordered,
                      name = "Vegetation class",
                      drop = FALSE, na.translate = FALSE) +
    labs(
      title = title_text,
      x = "Years since fire",
      y = NULL
    ) +
    #xlim(-1, 23) +
    theme_classic(base_size = 12) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 10)
    )

  plot_list[[i]] <- p

}
# 1. Extract the legend from a copy of plot 1 *with* legend
legend_plot <- plot_list[[1]] + theme(legend.position = "bottom", legend.direction = "horizontal")
legend <- get_legend(legend_plot)

# 2. Remove x-axis labels and ticks from all plots
plot_list <- lapply(plot_list, \(p) {
  p +
    theme(
      axis.title.x = element_blank(),
      legend.position = "none"
    )
})

# 3. Remove y-axis text from all but the first plot
plot_list[[2]] <- plot_list[[2]]
plot_list[[3]] <- plot_list[[3]] 

# 4. Combine the plots horizontally
plots_combined <- plot_grid(
  plotlist = plot_list,
  nrow = 1,
  labels = c("A", "B", "C"),
  label_size = 12,
  align = "v",
  axis = "l"
)

# 5. Shared axis labels
y_axis_label <- textGrob("Pixel count", gp = gpar(fontsize = 12), rot = 90)
x_axis_label <- textGrob("Years since fire", gp = gpar(fontsize = 12))

# 6. Stack: Y-label + Plots, then legend, then X-label
plot_with_y <- plot_grid(
  y_axis_label, plots_combined,
  rel_widths = c(0.04, 1),
  nrow = 1
)

final_plot <- plot_grid(
  plot_with_y,
  legend,
  x_axis_label,
  ncol = 1,
  rel_heights = c(1, 0.01, 0.07)
)

# 7. Show the final plot
print(final_plot)



```

#Map patch location w/in fire

```{r}
# Plot a fire map (outline-ish) with the patch in black
library(data.table)
library(ggplot2)
library(ggspatial)


#fire_id = c(10273, 10273, 10498),
#patch_id = c(3647, 3692, 3279)
#7514

# Load data
df <- fread("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/7514.csv")

# Plot with scale bar
fire_patch_map <- ggplot(df, aes(x = x, y = y)) +
  geom_point(color = "grey80", size = 0.4) +
  geom_point(data = df[id == 3902], color = "black", size = 0.4) +
  coord_fixed() +
  theme_void() +
  labs(title = "") +
  annotation_scale(location = "bl", width_hint = 0.2, 
                   text_cex = 0.6, line_width = 0.7)



```

```{r}

library(biscale)
library(scales)
library(ggspatial)

fire_id <- 7514
df_fire <- fread(paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/",fire_id,".csv"))

#find some large patches:

id_freq_rdnbr <- df_fire %>%
  group_by(id) %>%
  summarise(
    frequency = n(),
    rdnbr_class = first(rdnbr_class)  # or use `unique(rdnbr_class)` if paranoid about duplicates
  ) %>%
  arrange(desc(frequency))  # Optional: sort by frequency

print(id_freq_rdnbr)

patch_id <- 3538

df_patch <- df_fire %>%
  filter(id == patch_id)


# Pre-fire vegetation types within patch

veg_freq <- df_patch %>%
  group_by(RF_pre_veg) %>%
  summarise(
    frequency = n() 
  ) %>%
  arrange(desc(frequency))  # Optional: sort by frequency

print(veg_freq)


### Plot bivariate map returned within 20 yrs vs. distance from unburned greater/ less than 150m

# Define fire and patch IDs
fire_id <- 10094
patch_id <- 869

# Load data dynamically based on fire ID
file_path <- paste0("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/", fire_id, ".csv")
df_fire <- fread(file_path)

# Filter for specific patch and conifer vegetation
df_patch <- df_fire %>%
  filter(id == patch_id, RF_pre_veg == 7)

# Select necessary variables and remove NAs
df_map <- df_patch %>%
  select(x, y, returned, dist_unburned) %>%
  filter(!is.na(returned), !is.na(dist_unburned))





```

# Sankey plot of all fires pre-, 3 post and 20 yrs post-fire

```{r}
# Directory containing the CSV files
directory <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"

# Get a list of all CSV files in the directory
csv_files <- list.files(directory, pattern = "\\.csv$", full.names = TRUE)

# Filter files that contain any of the fire_IDs in their filename
filtered_files <- csv_files[sapply(csv_files, function(file) {
  fire_id <- as.numeric(str_extract(basename(file), "\\d+")) # Extract numeric ID from filename
  fire_id %in% megafire_IDs
})]

# Read and bind the filtered files into one dataframe
data <- bind_rows(lapply(filtered_files, fread))

data <- data %>%
  filter(rdnbr_class == 3)


library(networkD3)

# Step 1: Format the dataframe
df <- data %>%
  dplyr::select(RF_pre_veg, RF_post2_veg, RF_post20_veg) %>%
  mutate(across(everything(), as.character)) %>%
  filter(!is.na(RF_pre_veg), !is.na(RF_post2_veg), !is.na(RF_post20_veg))  # clean NAs

# Step 2: Add stage labels to differentiate node names
df <- df %>%
  mutate(
    pre = paste0(RF_pre_veg, " (pre)"),
    post2 = paste0(RF_post2_veg, " (2yr)"),
    post20 = paste0(RF_post20_veg, " (20yr)")
  )

# Step 3: Create links for two transitions
links_all <- bind_rows(
  df %>% count(source = pre, target = post2),
  df %>% count(source = post2, target = post20)
)

# Step 4: Create node list
nodes <- data.frame(name = unique(c(links_all$source, links_all$target)))

# Step 5: Add node IDs
nodes <- nodes %>%
  mutate(node_id = row_number() - 1)

# Step 6: Map source/target to IDs
links_all <- links_all %>%
  left_join(nodes, by = c("source" = "name")) %>%
  rename(source_id = node_id) %>%
  left_join(nodes, by = c("target" = "name")) %>%
  rename(target_id = node_id)

# Step 7: Define color palette (using numeric veg classes)
colors_all <- c(
  '1' = '#956733', '2' = '#a7b5a5', '3' = '#0940ca', '4' = '#d8bf58',
  '5' = '#6dcd2b', '6' = '#ffacc9', '7' = '#0e4f12', '8' = '#7b8d6a', '9' = '#e97451'
)

# Extract numeric part of node name to assign color
nodes$veg_code <- gsub(" .*", "", nodes$name)

# Create D3-compatible color scale
node_colors <- paste0(
  "d3.scaleOrdinal().domain([", 
  paste0("'", names(colors_all), "'", collapse = ", "), 
  "]).range([", 
  paste0("'", colors_all, "'", collapse = ", "), 
  "])"
)

# Step 8: Sankey plot
sankeyNetwork(
  Links = links_all,
  Nodes = nodes,
  Source = "source_id",
  Target = "target_id",
  Value = "n",
  NodeID = "name",
  NodeGroup = "veg_code",       # Color by veg class
  colourScale = node_colors,
  sinksRight = FALSE,
  fontSize = 12,
  nodeWidth = 30
)


```

## Counts for rdnbr_class/ veg combo by fire and overall

```{r}
library(data.table)
library(dplyr)
library(stringr)

# Directory and OBJECTID list
directory <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega"

# Initialize empty list to collect summary data
summary_list <- list()

# Loop over fire IDs
for (id in megafire_IDs) {
  file_path <- file.path(directory, paste0(id, ".csv"))
  
  if (file.exists(file_path)) {
    message("Processing OBJECTID: ", id)
    
    df <- tryCatch({
      # Read only relevant columns, disable memory mapping
      fread(file_path, select = c("rdnbr_class", "RF_pre_veg"), showProgress = FALSE)
    }, error = function(e) {
      warning(paste("Error reading file:", file_path, "-", e$message))
      return(NULL)
    })
    
    if (!is.null(df)) {
      # Summarize counts by group
      df_summary <- df %>%
        filter(!is.na(rdnbr_class), !is.na(RF_pre_veg)) %>%
        group_by(rdnbr_class, RF_pre_veg) %>%
        summarise(count = n(), .groups = "drop") %>%
        mutate(OBJECTID = id)
      
      summary_list[[as.character(id)]] <- df_summary
    }
  } else {
    warning("File not found for ID: ", id)
  }
}

# Combine into one master dataframe
master_summary <- bind_rows(summary_list)

# View first few rows
print(head(master_summary))


master_summary_all <- master_summary %>%
  group_by(rdnbr_class, RF_pre_veg) %>%
  summarise(total_count = sum(count), .groups = "drop")

master_summary_all$area_ha <- master_summary_all$total_count*0.09

fwrite(master_summary_all, "C:/Users/jscho/Documents/Megafires-1985-2023-Data/pre_veg_rdnbr_class_counts.csv")


```

## Veg time series stacked bars

```{r}
########################## veg classification time series #######################

library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

# Load the data
df <- fread("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/7808.csv")

# Define vegetation color palette (add "10" for 'Other')
veg_colors <- c(
  '7' = '#0e4f12',  # Conifer
  '1' = '#956733',
  '4' = '#d8bf58',
  '5' = '#6dcd2b',
  '8' = '#7b8d6a',
  '9' = '#e97451',
   '2'='#a7b5a5',
  '3'='#0940ca',
  '6'='#ffacc9'
)

# Step 1: Identify post-fire columns dynamically
veg_cols <- grep("^RF_post\\d+_veg$", names(df), value = TRUE)

# Step 2: Pivot to long format (post-fire)
df_post <- df %>%
  dplyr::select(all_of(veg_cols)) %>%
  mutate(row = row_number()) %>%
  pivot_longer(-row, names_to = "year", values_to = "veg_class") %>%
  mutate(
    year = as.integer(gsub("RF_post(\\d+)_veg", "\\1", year)),
    veg_class = as.character(veg_class)
  ) %>%
  group_by(year) %>%
  mutate(
    veg_class = ifelse(is.na(veg_class) & any(!is.na(veg_class)), "10", veg_class)  # replace NA if other values exist
  ) %>%
  ungroup() %>%
  filter(!is.na(veg_class)) %>%  # remove rows that are fully NA
  dplyr::select(-row)

# Step 3: Add RF_pre_veg as year -1
df_pre <- df %>%
  dplyr::select(RF_pre_veg) %>%
  rename(veg_class = RF_pre_veg) %>%
  mutate(
    year = -1,
    veg_class = as.character(veg_class)
  )

# Combine pre and post data
df_combined <- bind_rows(df_pre, df_post)

# Step 4: Calculate proportions
df_prop <- df_combined %>%
  group_by(year, veg_class) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(year) %>%
  mutate(proportion = count / sum(count))

# Set factor levels for stacking order: conifer ('7') at the bottom
df_prop$veg_class <- factor(df_prop$veg_class, levels = c("2","3", "4", "5","6", "8", "9","1","7"))

# Step 5: Plot
ggplot(df_prop, aes(x = year, y = proportion, fill = veg_class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_vline(xintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
  annotate("text", x = 1.8, y = 1.05, label = "*Fire", color = "red", fontface = "bold", size = 4) +
  scale_fill_manual(
    values = veg_colors,
    name = "Vegetation Class",
    labels = c(
      '1' = "Shrub",
      '4' = "Open woodland",
      '5' = "Herbaceous",
      '8' = "Sagebrush",
      '9' = "Dense woodland",
      '7' = "Conifer",
      '3' = "Water",
      '2' = "Bare rock",
      '6' = "Bare soil"
    )
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = sort(unique(df_prop$year))) +  # only show years that exist
  labs(
    title = "Post-fire Vegetation Composition (with Pre-fire)",
    x = "Years Since Fire",
    y = "Proportion of Area"
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )


# Combine plots with shared legend
combined_plot <- plot_grid(
  p1 + theme(legend.position = "none"),
  p2 + theme(legend.position = "none"),
  p3 + theme(legend.position = "none"),
  nrow = 1,
  labels = c("A", "B", "C"),
  label_size = 12
)

# Extract legend
legend <- get_legend(p1)

# Final panelled output
final_plot <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
print(final_plot)


```

# Split by planted/ not planted

```{r}
########################## veg classification time series #######################
# Rim fire (2013): 3983
# 3573 also long enough
# 8985
# Piute 5529
# Blue 7514

library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

# Load the data
df <- fread("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting/5529.csv")

# Define vegetation color palette (add "10" for 'Other')
veg_colors <- c(
  '7' = '#0e4f12',  # Conifer
  '1' = '#956733',
  '4' = '#d8bf58',
  '5' = '#6dcd2b',
  '8' = '#7b8d6a',
  '9' = '#e97451',
   '2'='#a7b5a5',
  '3'='#0940ca',
  '6'='#ffacc9'
)

# Step 1: Identify post-fire columns dynamically
veg_cols <- grep("^RF_post\\d+_veg$", names(df), value = TRUE)

# Step 2: Pivot to long format (post-fire)
df_post <- df %>%
  dplyr::select(all_of(veg_cols)) %>%
  mutate(row = row_number()) %>%
  pivot_longer(-row, names_to = "year", values_to = "veg_class") %>%
  mutate(
    year = as.integer(gsub("RF_post(\\d+)_veg", "\\1", year)),
    veg_class = as.character(veg_class)
  ) %>%
  group_by(year) %>%
  mutate(
    veg_class = ifelse(is.na(veg_class) & any(!is.na(veg_class)), "10", veg_class)  # replace NA if other values exist
  ) %>%
  ungroup() %>%
  filter(!is.na(veg_class)) %>%  # remove rows that are fully NA
  dplyr::select(-row)

# Step 3: Add RF_pre_veg as year -1
df_pre <- df %>%
  dplyr::select(RF_pre_veg) %>%
  rename(veg_class = RF_pre_veg) %>%
  mutate(
    year = -1,
    veg_class = as.character(veg_class)
  )

# Combine pre and post data
df_combined <- bind_rows(df_pre, df_post)

# Step 4: Calculate proportions
df_prop <- df_combined %>%
  group_by(year, veg_class) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(year) %>%
  mutate(proportion = count / sum(count))

# Set factor levels for stacking order: conifer ('7') at the bottom
df_prop$veg_class <- factor(df_prop$veg_class, levels = c("2","3", "4", "5","6", "8", "9","1","7"))

# Step 5: Plot
plot_veg_series_all <- ggplot(df_prop, aes(x = year, y = proportion, fill = veg_class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_vline(xintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
  annotate("text", x = 1.8, y = 1.05, label = "*Fire", color = "red", fontface = "bold", size = 4) +
  scale_fill_manual(
    values = veg_colors,
    name = "Vegetation Class",
    labels = c(
      '1' = "Shrub",
      '4' = "Open woodland",
      '5' = "Herbaceous",
      '8' = "Sagebrush",
      '9' = "Dense woodland",
      '7' = "Conifer",
      '3' = "Water",
      '2' = "Bare rock",
      '6' = "Bare soil"
    )
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = sort(unique(df_prop$year))) +  # only show years that exist
  labs(
    title = "Post-fire Vegetation Composition (with Pre-fire)",
    x = "Years Since Fire",
    y = "Proportion of Area"
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )


################ Split high severity into tree planting vs. non tree planting post-fire ##################

# Step 2: Long format with rdnbr_class
df_post <- df %>%
  dplyr::select(all_of(veg_cols), rdnbr_class, tree_planting) %>%
  mutate(row = row_number()) %>%
  pivot_longer(cols = -c(row, rdnbr_class, tree_planting), names_to = "year", values_to = "veg_class") %>%
  mutate(
    year = as.integer(gsub("RF_post(\\d+)_veg", "\\1", year)),
    veg_class = as.character(veg_class)
  ) %>%
  group_by(row, year, rdnbr_class, tree_planting) %>%
  mutate(
    veg_class = ifelse(is.na(veg_class) & any(!is.na(veg_class)), "10", veg_class)
  ) %>%
  ungroup() %>%
  filter(!is.na(veg_class)) %>%
  dplyr::select(-row)


# Step 3: Pre-fire vegetation (assume all share same rdnbr_class as in their row)
df_pre <- df %>%
  dplyr::select(RF_pre_veg, rdnbr_class, tree_planting) %>%
  rename(veg_class = RF_pre_veg) %>%
  mutate(
    year = -1,
    veg_class = as.character(veg_class)
  )


# Combine
df_combined <- bind_rows(df_pre, df_post)

# Step 4: Proportions
df_prop <- df_combined %>%
  group_by(rdnbr_class, year, tree_planting, veg_class) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(rdnbr_class, year, tree_planting) %>%
  mutate(proportion = count / sum(count))


# Step 5: Order vegetation class (stacking)
df_prop$veg_class <- factor(df_prop$veg_class,
                            levels = c("2", "3", "4", "5", "6", "8", "9", "1", "7"))


df_prop <- df_prop %>%
  mutate(
    rdnbr_class = as.character(rdnbr_class),
    rdnbr_class = dplyr::recode(rdnbr_class,
                         "0" = "Unchanged",
                         "1" = "Low",
                         "2" = "Moderate",
                         "3" = "High"),
    rdnbr_class = factor(rdnbr_class, levels = c("Unchanged", "Low", "Moderate", "High"))
  )


# Step 6: Plot with facets
df_plot <- df_prop

faceted_plot_planting <- ggplot(df_plot, aes(x = year, y = proportion, fill = veg_class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_vline(xintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
  annotate("text", x = 1.8, y = 1.05, label = "*Fire", color = "red", fontface = "bold", size = 4) +
  scale_fill_manual(
    values = veg_colors,
    name = "Vegetation class",
    labels = c(
      '1' = "Shrub",
      '4' = "Open woodland",
      '5' = "Herbaceous",
      '8' = "Sagebrush",
      '9' = "Dense woodland",
      '7' = "Conifer",
      '3' = "Water",
      '2' = "Bare rock",
      '6' = "Bare soil",
      '10' = "Other"
    )
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(
    title = "",
    x = "Years since fire",
    y = "Percent of total area"
  ) +
  facet_wrap(~tree_planting, labeller = as_labeller(c(`0` = "No Tree Planting", `1` = "Tree Planting"))) +
  theme_pubr() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )




################ Split by burn severity ##################################################################

df_plot <- df_prop %>% filter(rdnbr_class == "High")

faceted_plot_severity <- ggplot(df_plot, aes(x = year, y = proportion, fill = veg_class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_vline(xintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
  annotate("text", x = 1.8, y = 1.05, label = "*Fire", color = "red", fontface = "bold", size = 4) +
  scale_fill_manual(
    values = veg_colors,
    name = "Vegetation class",
    labels = c(
      '1' = "Shrub",
      '4' = "Open woodland",
      '5' = "Herbaceous",
      '8' = "Sagebrush",
      '9' = "Dense woodland",
      '7' = "Conifer",
      '3' = "Water",
      '2' = "Bare rock",
      '6' = "Bare soil",
      '10' = "Other"
    )
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(
    title = "",
    x = "Years since fire",
    y = "Percent of total area"
  ) +
  facet_wrap(~tree_planting, labeller = as_labeller(c(`0` = "No Tree Planting", `1` = "Tree Planting"))) +
  theme_pubr() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )

```

```{r}
########################## veg classification time series faceted by burn severity class #########################
# Size for exporting plots: 1050-830
# 7808- Storrie Fire 2000
# 9724- Campbell Fire 1990
# 7412- McNally Fire 2002
# 5718- Moonlight Fire 2007


# Load data
df <- fread("C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega/5718.csv")

# Define color palette
veg_colors <- c(
  '7' = '#0e4f12',  # Conifer
  '1' = '#956733',
  '4' = '#d8bf58',
  '5' = '#6dcd2b',
  '8' = '#7b8d6a',
  '9' = '#e97451',
  '2' = '#a7b5a5',
  '3' = '#0940ca',
  '6' = '#ffacc9'
)

# Step 1: Identify post-fire columns
veg_cols <- grep("^RF_post\\d+_veg$", names(df), value = TRUE)

# Step 2: Long format with rdnbr_class
df_post <- df %>%
  select(all_of(veg_cols), rdnbr_class) %>%
  mutate(row = row_number()) %>%
  pivot_longer(cols = -c(row, rdnbr_class), names_to = "year", values_to = "veg_class") %>%
  mutate(
    year = as.integer(gsub("RF_post(\\d+)_veg", "\\1", year)),
    veg_class = as.character(veg_class)
  ) %>%
  group_by(row, year, rdnbr_class) %>%
  mutate(
    veg_class = ifelse(is.na(veg_class) & any(!is.na(veg_class)), "10", veg_class)
  ) %>%
  ungroup() %>%
  filter(!is.na(veg_class)) %>%
  select(-row)

# Step 3: Pre-fire vegetation (assume all share same rdnbr_class as in their row)
df_pre <- df %>%
  select(RF_pre_veg, rdnbr_class) %>%
  rename(veg_class = RF_pre_veg) %>%
  mutate(
    year = -1,
    veg_class = as.character(veg_class)
  )

# Combine
df_combined <- bind_rows(df_pre, df_post)

# Step 4: Proportions
df_prop <- df_combined %>%
  group_by(rdnbr_class, year, veg_class) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(rdnbr_class, year) %>%
  mutate(proportion = count / sum(count))

# Step 5: Order vegetation class (stacking)
df_prop$veg_class <- factor(df_prop$veg_class,
                            levels = c("2", "3", "4", "5", "6", "8", "9", "1", "7"))


df_prop <- df_prop %>%
  mutate(
    rdnbr_class = as.character(rdnbr_class),  # ensure character for recode
    rdnbr_class = dplyr::recode(rdnbr_class,
                         "0" = "Unchanged",
                         "1" = "Low",
                         "2" = "Moderate",
                         "3" = "High"),
    rdnbr_class = factor(rdnbr_class, levels = c("Unchanged", "Low", "Moderate", "High"))
  )

# Step 6: Plot with facets
ggplot(df_prop, aes(x = year, y = proportion, fill = veg_class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_vline(xintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
  annotate("text", x = 1.8, y = 1.05, label = "*Fire", color = "red", fontface = "bold", size = 4) +
  scale_fill_manual(
    values = veg_colors,
    name = "Vegetation class",
    labels = c(
      '1' = "Shrub",
      '4' = "Open woodland",
      '5' = "Herbaceous",
      '8' = "Sagebrush",
      '9' = "Dense woodland",
      '7' = "Conifer",
      '3' = "Water",
      '2' = "Bare rock",
      '6' = "Bare soil",
      '10' = "Other"
    )
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(
    title = "",
    x = "Years since fire",
    y = "Percent of total area"
  ) +
  facet_wrap(~rdnbr_class) +
  theme_pubr() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )
```

# Megafire table

```{r}
# Load every csv file, calculate summary metrics per fire and save in master table

library(data.table)
library(fs)

# Set input directory
input_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting/"

# List CSV files
csv_files <- dir_ls(input_dir, glob = "*.csv")

# Initialize empty list to hold summaries
summary_list <- list()

# Loop through each CSV
for (csv_file in csv_files) {
  message("Processing: ", path_file(csv_file))
  
  # Read data
  df <- fread(csv_file)
  
  # Skip if required columns are missing
  if (!all(c("OBJECTID", "fire_year", "rdnbr_class", "RF_pre_veg") %in% names(df))) {
    warning("Missing required columns in: ", csv_file)
    next
  }
  
  # Extract fire metadata
  fire_id <- unique(df$OBJECTID)
  fire_year <- unique(df$fire_year)
  
  if (length(fire_id) != 1 || length(fire_year) != 1) {
    warning("Inconsistent OBJECTID or fire_year in: ", csv_file)
    next
  }
  
  # Count rdnbr_class values (0 to 3)
  rdnbr_counts <- as.list(table(factor(df$rdnbr_class, levels = 0:3)))
  names(rdnbr_counts) <- paste0("rdnbr_", names(rdnbr_counts))
  
  # Count RF_pre_veg values (1 to 9)
  veg_counts <- as.list(table(factor(df$RF_pre_veg, levels = 1:9)))
  names(veg_counts) <- paste0("veg_", names(veg_counts))
  
  # Combine everything
  summary_row <- c(
    OBJECTID = fire_id,
    fire_year = fire_year,
    rdnbr_counts,
    veg_counts
  )
  
  summary_list[[length(summary_list) + 1]] <- summary_row
}

# Combine into a master data.table
summary_dt <- rbindlist(summary_list, fill = TRUE)

# Save if needed
# fwrite(summary_dt, "C:/Users/jscho/Documents/Megafires-1985-2023-Data/summary_metrics_by_fire.csv")

# Show the summary
summary_dt

# ---- 🔗 Join Fire Name, Area, and Cause from fire_metrics_mega ----
summary_dt[, OBJECTID := as.integer(OBJECTID)]

summary_dt <- summary_dt %>%
  left_join(
    fire_metrics_mega %>%
      dplyr::select(OBJECTID, FIRE_NAME, area_ha, CAUSE) %>%
      dplyr::rename(Fire = FIRE_NAME),
    by = "OBJECTID"
  ) %>%
  select(Fire, area_ha, CAUSE, everything(), -OBJECTID)

# ---- 🛠 Recode CAUSE values ----
# First, make sure CAUSE is character
summary_dt[, CAUSE := as.character(CAUSE)]

# Define recode mapping
cause_labels <- c(
  "1" = "Lightning",
  "2" = "Equipment use",
  "3" = "Smoking",
  "4" = "Campfire",
  "5" = "Debris",
  "6" = "Railroad",
  "7" = "Arson",
  "8" = "Playing with fire",
  "9" = "Miscellaneous",
  "10" = "Vehicle",
  "11" = "Powerline",
  "12" = "Firefighter training",
  "13" = "Non-firefighter training",
  "14" = "Unknown/ unidentified",
  "15" = "Structure",
  "16" = "Aircraft",
  "17" = "Volcanic",
  "18" = "Escaped prescribed burn",
  "19" = "Illegal alien campfire"
)

# Recode CAUSE using the mapping
summary_dt[, CAUSE := cause_labels[CAUSE]]



summary_dt <- summary_dt %>%
  mutate(area_kha = round(area_ha / 1000, 1)) %>%
  select(Fire, area_kha, CAUSE, everything(), -area_ha)

fwrite(summary_dt, "C:/Users/jscho/Documents/Megafires-1985-2023-Data/summary_metrics_by_fire.csv")
```

```{r}
### Calculate some areas
library(sf)


### Area within megafires on FS land with tree plantings

# Load the shapefile
shp <- st_read("C:/Users/jscho/Documents/Megafires-1985-2023-Data/Shapefiles/FACTS_tree_planting_1986_2024_in_megafires.shp")

# Ensure the CRS is projected (not in degrees)
st_crs(shp)


# If needed, reproject to an appropriate projected CRS (e.g., EPSG:3310 for California)
shp <- st_transform(shp, crs = 3310)

# Calculate areas in square meters
areas_m2 <- st_area(shp)

# Total area (automatically returns units in m²)
total_area_m2 <- sum(areas_m2)

# Convert to hectares
total_area_ha <- units::set_units(total_area_m2, ha)

total_area_plantings_ha <- total_area_ha

# Areas within megafires on FS land total
# Load the shapefile
shp <- st_read("C:/Users/jscho/Documents/Megafires-1985-2023-Data/Shapefiles/megafires_in_NF_area.shp")

# Ensure the CRS is projected (not in degrees)
st_crs(shp)


# If needed, reproject to an appropriate projected CRS (e.g., EPSG:3310 for California)
shp <- st_transform(shp, crs = 3310)

# Calculate areas in square meters
areas_m2 <- st_area(shp)

# Total area (automatically returns units in m²)
total_area_m2 <- sum(areas_m2)

# Convert to hectares
total_area_ha <- units::set_units(total_area_m2, ha)

total_area_FS_ha <- total_area_ha


# Total megafire area
# Load the shapefile
shp <- st_read("C:/Users/jscho/Documents/Megafires-1985-2023-Data/Shapefiles/SN_megafires_1985_2023.shp")

# Ensure the CRS is projected (not in degrees)
st_crs(shp)


# If needed, reproject to an appropriate projected CRS (e.g., EPSG:3310 for California)
shp <- st_transform(shp, crs = 3310)

# Calculate areas in square meters
areas_m2 <- st_area(shp)

# Total area (automatically returns units in m²)
total_area_m2 <- sum(areas_m2)

# Convert to hectares
total_area_ha <- units::set_units(total_area_m2, ha)

total_area_megafires_ha <- total_area_ha

```

# ANOVA NEW

```{r}
# Needs cleaning up!


#Fire level datasets
fire_metrics_mega <- fread("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Dataframes/Megafire_paper/SN_megafires_1985_2023.csv", header=TRUE)

megafire_IDs <- c(5529, 5718, 5721, 8026, 7808, 7747, 7996, 7961, 
                  7412, 7514, 7242, 9714, 9724, 9487, 8985, 9039, 
                  8747, 10495, 10498, 10504, 10094, 10273, 1819, 1772, 
                  2111, 2140, 1210, 1211, 1216, 1215, 1220, 1319, 
                  1315, 718, 911, 932, 869, 888, 945, 959, 504, 
                  3363, 3567, 3573, 3574, 3051, 3094, 3188, 2706, 
                  2147, 2531, 2415, 2618, 5207, 5309, 4277, 4285, 
                  4306, 4380, 4386, 4374, 4500, 3983, 3992)

#Patch level datasets- rdnbr
patch_metrics_mega <- fread("C:/Users/jscho/OneDrive - University of Cambridge/PhD/Analyses/Dataframes/Megafire_paper/SN_patch_metrics_mega_rdnbr_1985_2023.csv")

# Fire IDs
fire_IDs <- c(10495,10498,10504,10273,10094,9714,9724,9487,8985,
              9039,8747,8026,7996,7961,7808,7747,7514,7412,7242)

# File paths
csv_dir <- "C:/Users/jscho/Documents/Megafires-1985-2023-Data/raster_df_mega_repeated_planting"
csv_files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[basename(csv_files) %in% paste0(fire_IDs, ".csv")]


### Excluding patches with any pixels reburned or planted ###

# Initialize list to hold per-fire summaries
patch_list <- list()

for (file in csv_files) {
  fire_id <- as.integer(gsub("\\.csv$", "", basename(file)))
  dt_full <- fread(file)
  
  # --- 🧱 Step 1: Pre-filter out all patches that ever had planting or reburn ---
  excluded_ids <- unique(dt_full[tree_planting == 1 | reburn_20yrs == 1, id])
  dt_full <- dt_full[!id %in% excluded_ids]
  
  # --- 📦 Step 2: Compute patch size & rdnbr_class from all pixels ---
  patch_sizes <- dt_full[, .(n_total = .N, rdnbr_class = first(rdnbr_class)), by = id]
  
  # --- 🧼 Step 3: Now filter for pixels of interest ---
  dt <- dt_full[RF_pre_veg == 7 & transitioned == 1]
  
  if (nrow(dt) == 0) next
  
  # Add flag for 20-year recovery
  dt[, returned_20yrs := as.integer(returned == 1 & yrs_to_return < 21)]
  
  # Merge patch size info
  dt <- merge(dt, patch_sizes, by = "id", all.x = TRUE)
  dt[, rdnbr_class := rdnbr_class.x]
  dt[, rdnbr_class.x := NULL]
  dt[, rdnbr_class.y := NULL]
  
  
  # --- 🧾 Step 4: Summarize at patch level ---
  patch_summary <- dt[, .(
    fire_id = fire_id,
    rdnbr_class = first(rdnbr_class),  # from patch_sizes
    n_total = first(n_total),
    n_returned = sum(returned_20yrs, na.rm = TRUE)
  ), by = id]
  
  # Assign patch size class
  patch_summary[, patch_size_class := cut(
    n_total,
    breaks = c(1, 100, 1000, 10000, 1500000),
    labels = c("small", "medium", "large", "very_large"),
    include.lowest = TRUE
  )]
  
  patch_summary[, abs_not_recovered := n_total - n_returned]
  
  patch_list[[length(patch_list) + 1]] <- patch_summary
}

# Combine all fires' patch summaries
patch_data_clean <- rbindlist(patch_list)

# Filter out any NA patch size or missing rdnbr_class
patch_data_clean <- patch_data_clean[!is.na(patch_size_class) & !is.na(rdnbr_class) & rdnbr_class > 0]


patch_data_clean <- patch_data_clean %>%
  mutate(n_bins = cut(n_total,
                      breaks = c(1, 100, 1000,10000,1500000),
                      labels = c("small","medium","large","very_large"),
                      include.lowest = TRUE))



patch_data_clean$prop_orig_con_20 <- patch_data_clean$n_returned / patch_data_clean$n_total



### Run ANOVA ###
# ANOVA
patch_data_clean <- patch_data_clean %>%
  mutate(n_bins_rdnbr = paste0(as.character(n_bins), "_", as.character(rdnbr_class)))

patch_data_clean$rdnbr_class <- as.factor(patch_data_clean$rdnbr_class)

# Run one-way ANOVA- This tests if mean prop_orig_con_20 differs by group (i.e., small_1, medium_3, etc.).
anova_result <- aov(prop_orig_con_20 ~ n_bins_rdnbr, data = patch_data_clean)
summary(anova_result)


# Two-way ANOVA with interaction- This will show Main effects of patch size (n_bins), Main effects of burn severity (rdnbr_class), Interaction effects between the two

anova_interaction <- aov(prop_orig_con_20 ~ n_bins * rdnbr_class, data = patch_data_clean)
summary(anova_interaction)

# Normality of residuals
plot(anova_result, which = 2)  # Q-Q plot

# Homogeneity of variance
plot(anova_result, which = 1)  # Residuals vs fitted

# Levene's Test (optional, from car package)
library(car)
leveneTest(prop_orig_con_20 ~ n_bins_rdnbr, data = patch_data_clean)

TukeyHSD(anova_result)

TukeyHSD(anova_interaction)





### Original: with patch_metrics dataframe ######
# For 20 year post-fire series: patch size vs. perc. recov after 20 years

patch_metrics_mega <- patch_metrics_mega %>%
  mutate(n_bins = cut(n,
                      breaks = c(1, 100, 1000,10000,1500000),
                      labels = c("small","medium","large","very_large"),
                      include.lowest = TRUE))

# Select patches with at least 20 years post-fire history
filtered_patch_metrics <- patch_metrics_mega %>%
  filter(OBJECTID %in% megafire_IDs)

filtered_patch_metrics$prop_con_trans <- 1- (filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7)
filtered_patch_metrics$prop_orig_con_1 <- filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_3 <- filtered_patch_metrics$count_RF_post1_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_5 <- filtered_patch_metrics$count_RF_post5_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_10 <- filtered_patch_metrics$count_RF_post10_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_15 <- filtered_patch_metrics$count_RF_post15_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$prop_orig_con_20 <- filtered_patch_metrics$count_RF_post20_veg_7 / filtered_patch_metrics$count_RF_pre_veg_7
filtered_patch_metrics$abs_con_notRecov_20yrs <- filtered_patch_metrics$count_RF_pre_veg_7 - filtered_patch_metrics$count_RF_post20_veg_7


filtered_patch_metrics_HS <- filtered_patch_metrics %>%
  filter(rdnbr_class==3)

filtered_patch_metrics_5con <- filtered_patch_metrics %>%
  filter(count_RF_pre_veg_7 > 4)

filtered_patch_metrics_HS_5con <- filtered_patch_metrics_HS %>%
  filter(count_RF_pre_veg_7 > 4)

filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  filter(rdnbr_class >0)

# Combine n_bins and rdnbr_class into a single factor for the x-axis
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = interaction(n_bins, rdnbr_class, sep = "_"))


# Add spacing to the groups by adding empty levels
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = as.factor(n_bins_rdnbr)) %>%
  arrange(n_bins_rdnbr, rdnbr_class) %>%
  mutate(
    n_bins_rdnbr = factor(n_bins_rdnbr, levels = c(levels(n_bins_rdnbr), "")),  # Add an empty level for gaps
    rdnbr_class = as.factor(rdnbr_class)  # Ensure rdnbr_class is a factor
  )

# Summarize data for the marginal bar plot
marginal_data <- filtered_patch_metrics_5con %>%
  group_by(n_bins_rdnbr) %>%
  summarize(total_abs_con_notRecov = sum(abs_con_notRecov_20yrs, na.rm = TRUE))

marginal_data$total_abs_con_notRecov_ha <- marginal_data$total_abs_con_notRecov * 0.09

# Inspect unique values in n_bins_rdnbr
unique_bins <- unique(filtered_patch_metrics_5con$n_bins_rdnbr)

# Reorder levels dynamically based on observed values
filtered_patch_metrics_5con <- filtered_patch_metrics_5con %>%
  mutate(n_bins_rdnbr = factor(n_bins_rdnbr, levels = c("small_1","small_2","small_3", "medium_1","medium_2","medium_3", "large_1","large_2","large_3","very_large_1","very_large_2","very_large_3")))  # Replace with actual levels if different


# ANOVA

# Run one-way ANOVA- This tests if mean prop_orig_con_20 differs by group (i.e., small_1, medium_3, etc.).
anova_result <- aov(prop_orig_con_20 ~ n_bins_rdnbr, data = filtered_patch_metrics_5con)
summary(anova_result)


# Two-way ANOVA with interaction- This will show Main effects of patch size (n_bins), Main effects of burn severity (rdnbr_class), Interaction effects between the two

anova_interaction <- aov(prop_orig_con_20 ~ n_bins * rdnbr_class, data = filtered_patch_metrics_5con)
summary(anova_interaction)

# Normality of residuals
plot(anova_result, which = 2)  # Q-Q plot

# Homogeneity of variance
plot(anova_result, which = 1)  # Residuals vs fitted

# Levene's Test (optional, from car package)
library(car)
leveneTest(prop_orig_con_20 ~ n_bins_rdnbr, data = filtered_patch_metrics_5con)

TukeyHSD(anova_result)

TukeyHSD(anova_interaction)

```
