Code by Johanna Schönecker

Edited 15 October 2025

This quarto document contains code to load and pre-process pixel-, fire- and patch-level data for megafires in the Sierra Nevada ecoregion which burned between 1985 and 2023.

Different sections contain code to calculate area weighted mean patch sizes (AWMPS) for each megafire/ burn severity class as well as the largest patch index (LPI) for each megafire/ burn severity class

##Packages

```{r Loading packages}
# -------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyr, raster, rgdal, ggplot2, ggspatial, sp, sf, rgeos, reshape,
  patchwork, Rcpp, scales, dplyr, ggpubr, landscapemetrics, landscapetools,
  here, stringr, moments, reshape2, ggeasy, vctrs, mgcv, networkD3,
  ggridges, cowplot, factoextra, gdistance, fs, data.table, terra,
  gridExtra, grid, car, biscale, e1071, scales, devEMF
)

```

##Data loading and prep

```{r section-chunk2}
#Fire level datasets
fire_metrics_mega <- fread(paste0(here(),"/Data/SN_megafires_1985_2023.csv"), header=TRUE)

#Patch level datasets- rdnbr
patch_metrics_mega <- fread(paste0(here(),"/Data/SN_patch_metrics_mega_rdnbr_1985_2023.csv"))

megafire_IDs <- c(5529, 5718, 5721, 8026, 7808, 7747, 7996, 7961, 
                   7412, 7514, 7242, 9714, 9724, 9487, 8985, 9039, 
                   8747, 10495, 10498, 10504, 10094, 10273, 1819, 1772, 
                   2111, 2140, 1210, 1211, 1216, 1215, 1220, 1319, 
                   1315, 718, 911, 932, 869, 888, 945, 959, 504, 
                   3363, 3567, 3573, 3574, 3051, 3094, 3188, 2706, 
                   2147, 2531, 2415, 2618, 5207, 5309, 4277, 4285, 
                   4306, 4380, 4386, 4374, 4500, 3983, 3992)
```


#AWMPS- area weighted mean patch size

```{r section-chunk6}
# -------------------------------------------------------------------
# Function to calculate AWMPS for a given set of patch areas
# -------------------------------------------------------------------

calculate_awmps <- function(patch_areas) {
  total_area <- sum(patch_areas)
  awmps <- sum(patch_areas^2) / total_area
  return(awmps)
}

# -------------------------------------------------------------------
# Create empty dataframe to store results
# -------------------------------------------------------------------
patch_size_aw_mega <- data.frame(
  rdnbr_class = integer(),
  AWMPS       = numeric(),
  OBJECTID    = integer()
)

# -------------------------------------------------------------------
# Loop through each megafire ID and calculate AWMPS
# -------------------------------------------------------------------
for (i in megafire_IDs) {
  
  patch_metrics_indiv <- patch_metrics_mega %>%
    filter(OBJECTID == i)
  
  # Calculate AWMPS for each severity class
  awmps_by_class <- patch_metrics_indiv %>%
    group_by(rdnbr_class) %>%
    summarise(AWMPS = calculate_awmps(n), .groups = "drop")
  
  awmps_by_class$OBJECTID <- i
  
  patch_size_aw_mega <- rbind(patch_size_aw_mega, awmps_by_class)
}

# -------------------------------------------------------------------
# Clean and recode severity classes
# -------------------------------------------------------------------
patch_size_aw_mega <- patch_size_aw_mega %>%
  filter(rdnbr_class != -1) %>%
  mutate(
    rdnbr_class = as.character(rdnbr_class),
    rdnbr_class = dplyr::recode(
      rdnbr_class,
      "1" = "low",
      "2" = "moderate",
      "3" = "severe"
    ),
    awmps_ha = AWMPS * 0.09
  )


```

# LPI

```{r lpi-chunk1}
# -------------------------------------------------------------------
# Calculate the LPI for every class in every megafire
# -------------------------------------------------------------------

# Path to classified RdNBR rasters
raster_dir <- paste0(here(),"/Data/Rasters/RdNBR_classified")

# List all .tif raster files in the folder
raster_files <- fs::dir_ls(raster_dir, regexp = "\\.tif$")

# Function to calculate LPI per class for a raster
calculate_lpi <- function(file_path) {
  # Load raster
  rast_obj <- terra::rast(file_path)

  # Convert to RasterLayer (landscapemetrics prefers Raster*)
  rast_r <- raster::raster(rast_obj)

  # Calculate LPI for each class
  lpi_result <- landscapemetrics::lsm_c_lpi(rast_r)

  # Add raster name to identify the source
  lpi_result$raster_name <- basename(file_path)

  lpi_result
}

# Apply to all rasters and combine
lpi_list <- lapply(raster_files, calculate_lpi)
lpi_df   <- dplyr::bind_rows(lpi_list)

# -------------------------------------------------------------------
# Prepare a wide table for supplement
# -------------------------------------------------------------------

# Step 1: Extract fire ID from raster_name and join fire names
lpi_df <- lpi_df %>%
  dplyr::mutate(
    OBJECTID = as.numeric(stringr::str_remove(raster_name, "\\.tif$"))
  ) %>%
  dplyr::left_join(
    fire_metrics_mega %>% dplyr::select(OBJECTID, FIRE_NAME),
    by = "OBJECTID"
  )

# Step 2: Recode class into labels
lpi_df <- lpi_df %>%
  dplyr::mutate(
    class_label = dplyr::case_when(
      class == 0 ~ "Unchanged",
      class == 1 ~ "Low",
      class == 2 ~ "Moderate",
      class == 3 ~ "High",
      TRUE       ~ as.character(class)
    )
  )

# Step 3: Summarise to one row per fire with total value per class
lpi_summary <- lpi_df %>%
  dplyr::group_by(FIRE_NAME, class_label) %>%
  dplyr::summarise(total_value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from  = class_label,
    values_from = total_value
  ) %>%
  dplyr::rename(`Fire name` = FIRE_NAME) %>%
  dplyr::select(`Fire name`, Unchanged, Low, Moderate, High) %>%
  dplyr::left_join(
    fire_metrics_mega %>% dplyr::select(FIRE_NAME, YEAR_),
    by = c("Fire name" = "FIRE_NAME")
  ) %>%
  dplyr::rename(`Fire year` = YEAR_) %>%
  dplyr::select(`Fire name`, `Fire year`, Unchanged, Low, Moderate, High)

# Write CSV
data.table::fwrite(
  lpi_summary, paste0(here(),"/Outputs/LPI_all_megafires.csv")
)


```
